A retrospective analysis of 400 publications reveals patterns of irreproducibility across an entire life sciences research field

Version 1

Abstract

The ReproSci project retrospectively analyzed the reproducibility of 1006 claims from 400 papers published between 1959 and 2011 in the field of Drosophila immunity. This project attempts to provide a comprehensive assessment, 14 years later, of the replicability of nearly all publications across an entire scientific community in experimental life sciences. We found that 61% of claims were verified, while only 7% were directly challenged (not reproducible), a replicability rate higher than previous assessments. Notably, 24% of claims had never been independently tested and remain unchallenged. We performed experimental validations of a selection of 45 unchallenged claim, that revealed that a significant fraction (38/45) of them is in fact non-reproducible. We also found that high-impact journals and top-ranked institutions are more likely to publish challenged claims. In line with the reproducibility crisis narrative, the rates of both challenged and unchallenged claims increased over time, especially as the field gained popularity. We characterized the uneven distribution of irreproducibility among first and last authors. Surprisingly, irreproducibility rates were similar between PhD students and postdocs, and did not decrease with experience or publication count. However, group leaders, who had prior experience as first authors in another Drosophila immunity team, had lower irreproducibility rates, underscoring the importance of early-career training. Finally, authors with a more exploratory, short-term engagement with the field exhibited slightly higher rates of challenged claims and a markedly higher proportion of unchallenged ones. This systematic, field-wide retrospective study offers meaningful insights into the ongoing discussion on reproducibility in experimental life sciences.

Introduction

Science predominantly thrives on the continuous aggregation of knowledge, with researchers progressively building upon one another’s discoveries. As results are replicated (when independent studies reach the same conclusions as the original authors of a claim), confidence in the evidence supporting the claim grows. Confidence in publishing results is critical to sustain trust within a scientific community and to allow new research to build on previous findings (Fidler and Wilcox, 2018). Over the past decades, this confidence has been on a decline.

Several recent reports point out that reproducibility is lower than desirable in molecular life sciences (Amaral et al., 2025; Begley and Ioannidis, 2015; Fidler et al., 2017; Ioannidis et al., 2009; Macleod et al., 2014). In 2016, an extensive survey indicated that more than 70% of researchers have tried and failed to reproduce another scientist’s experiments, and more than half have failed to reproduce their own experiments (Baker, 2016). Another study suggested that about 85% of biomedical research is ‘wasted’, meaning that inefficiencies and methodological flaws throughout the research process undermine the value of scientific findings (Begley and Ioannidis, 2015; Chalmers and Glasziou, 2009). In addition, 52% of scientists that were surveyed by Nature agree that we are currently facing a significant ‘crisis’ of replicability. Ongoing concerns about replicability in academic research carry important consequences for the private sector. Two pharmaceutical companies, Bayer and Amgen, reported replicability rates of 11% and 25% in two independent efforts to reproduce findings from dozens of groundbreaking basic studies in cancer and related areas (Prinz et al., 2011). Moreover, an extensive reproducibility project in cancer biology reveals that a very significant fraction of findings reported in high profile journals cannot be replicated or have smaller effect size that initially reported (Errington et al., 2021, 2014). Lack of replicability can have multiple causes, ranging from the presence cryptic factors not accounted for by authors, misuse of technical tools, spoiled reagents, bias in data reporting, lack of statistical power or to even outright fraud (Devezer and Buzbas, 2023; Eisner, 2018; Lesperance and Broderick, 2021; Peng, 2015).

Misleading or erroneous findings can disrupt scientific progress, wasting valuable time and resources as researchers attempt to replicate or use the results of a published article (Stern et al., 2014). In some cases, entire research initiatives may be built upon findings that are later discovered to be inaccurate or flawed (Piller, 2022). This might also result in unfair career outcomes if sensational but irreproducible results are rewarded over rigorous studies, especially in high-impact publications (Lemaitre, 2016; Steneck, 2011). However, zero irreproducibility is neither realistic nor necessarily desirable. A certain level of irreproducibility may reflect scientific boldness and risk-taking. Cutting-edge explorations of the unknown are expected to involve occasional backtracking as the bigger picture gradually comes into focus. Thus, striving for perfect replicability could paradoxically stifle innovation and promote overly conservative science. Determining the ideal balance is currently an open question (Shiffrin et al., 2018).

Considering the prominent role played by science in contemporary society, there is a growing need to understand how academic incentives influence the reliability of published results. Indeed, the field of metascience - the scientific investigation of science itself - is gaining popularity (Fortunato et al., 2018). Analyzing how replicability affects the scientific enterprise is required to better characterize how science progresses despite mistakes. It can also help to influence policy to improve science replicability, or at least to limit its most detrimental impacts.

Research focusing on the immune system of Drosophila flourished in the early 1990s and is still an active field of investigation. It boasts two thousand articles published since its beginning, dealing with various facets of the immune response of this insect, such as humoral immunity, cellular immunity, response to pathogens, immuno-metabolism, and evolution of the immune system (Westlake et al., 2024). While the field remains modest in size, it gained visibility in mid 1990s when researchers realized that some mechanisms regulating the Drosophila immune defense are conserved in mammals, including humans. The discovery of Toll-like receptors, stemming from the identification of Toll’s role in the Drosophila immune response (Lemaitre, 2004), established the value and utility of the field to immunological research, a finding that was acknowledged by a Nobel prize awarded to Jules Hoffmann in 2011.

The medium size of this research field and the lack of direct biomedical influence are ideal for a metascience analysis. The Drosophila immunity field is substantially smaller than other fields of research such as HIV, microbiota or cancer, where the rapid accumulation of articles forbids any comprehensive evaluation of scientific claims by an individual or a laboratory. With fewer direct translational or biomedical influences, we expect scientists working on Drosophila immunity to be less affected by external biases, and to prioritize recognition by peers. Moreover, Drosophila research is generally perceived as solid, because it relies predominantly on its powerful genetics and well-controlled experimental conditions, so the field is particularly interesting for this analysis.

In the framework of the ReproSci project, we retrospectively analyzed the replication of 400 articles published between 1959 and 2011 in the field of Drosophila immunity research-that is, nearly all the major articles published by a community of scientists in experimental life science over 50 years (Westlake et al., 2025). For each article, we identified the main, major, and minor scientific conclusions, hereafter referred to as “claims”. The 14-year interval between the time the latest article was published and our study, during which scientific knowledge has advanced and other laboratories have had the opportunity to attempt replication, provides the necessary perspective to efficiently assess the validity of the claims. In a first step, we created a publicly accessible website, displaying all the information on the articles, the claims, and their validity (https://ReproSci.epfl.ch/). Second, we contacted the Drosophila scientific community to provide feedback via this website on the replicability of the claims. In a third step, several teams attempted to replicate 45 major and 11 minor claims that remained unchallenged in the literature at the time. The present article focuses on the meta-science findings, while the scientific consequences of the ReproSci project in the Drosophila immunity field are presented in a companion article (Westlake et al., 2025).

This article analyzes the pattern of replicability with two objectives: first, to comprehend the level of replicability within the experimental life science community, and second, to examine whether instances of non-replicability tend to aggregate among journals, first authors, and lead researchers. Such clustering may identify individual characteristics and environmental factors that influence science replicability. We open the Results section with a detailed exploratory analysis of the effect of individual categorical variables on reproducibility outcomes. Then, to assess the joint influence of all covariates and account for the clustering of multiple claims by the same author, we fit a multivariate model to account for confounding effects across our covariates. Because challenged claims are rare and the overall sample is modest, statistical power remains limited. However, taken together, i) the exploratory and ii) multivariate analyses form a single narrative, the first describing exhaustively the observed patterns, and the second testing the robustness of the associations.

Results

Methodology and claim assessment

Scientific claims were divided into three categories. The main (one per article), major (3-4) and minor (2-4) claims were manually extracted from 400 articles in the field of Drosophila immunity published before 2011. We then assessed the claims of each article to determine if they were supported or refuted by subsequent publications. The claims were manually categorized by H.W. and B.L. into four broad groups: i) verified by subsequent studies (when later studies confirmed the claim), ii) challenged (when later studies contradicted or disproved the claim), iii) unchallenged (when no later study addressed or attempted to replicate the claim), and iv) partially verified or mixed (when later studies did not lead to a definitive conclusion (see Table 1 in Methods for classification and Westlake et al., 2025). Then, experimental work was performed in several laboratories to test the validity of 45 major and 11 minor unchallenged claims using alternate methods. These claims subjected to experimental validation were selected according to criteria such as appearing suspicious due to the absence of direct follow up or being straightforward to test experimentally. Our data and findings were put on a community-accessible website that opened to the public in early July 2023. The present article focusses on the pattern of irreproducibility and statistical analyses were performed on the 1006 major claims only (referred to as claims below) of the database as updated on December 20th 2024. In this article, challenged claims will interchangeably be called irreproducible. Our project focuses on assessing the strength of the claims themselves (i.e. indirect reproducibility or conceptual replicability) rather than testing whether the original methods produce repeatable results (direct reproducibility/replicability). Thus, our conclusions do not directly challenge the initial results leading to a claim, but rather the general applicability (i.e. generalizability) of the claim itself (Goodman et al., 2016; Meng, 2020).

Drosophila immunity claims are mostly replicable

After an extensive literature review to assess the validity of claims published from 1959 to 2011, combined with experimental testing of 45 major claims, we found that approximately 61% of major claims (610 out of 1006 across the 400 selected articles) were fully verified (Table 2). Among these 610 verified claims, 91.7% were corroborated by independent laboratories, 7.2% were confirmed by the same laboratory, and 0.7% were verified through experimental work conducted as part of the ReproSci project. This verification rate is higher than that observed in other fields where reproducibility has been systematically assessed (Amaral et al., 2025; Amaral et al., 2019; Da Costa et al., 2022; Errington et al., 2021; Fidler et al., 2017; Youyou et al., 2023). These results may reflect the robust scientific standards and methodological rigor of Drosophila research, which benefits from the availability of diverse and powerful genetics approaches and from a collaborative culture of rapid reagent sharing, that facilitates replication across laboratories.

We also observed that 23.8% of major claims had no published follow-up (Table 2, Figure 1); these unchallenged claims are particularly interesting and will be discussed in detail in the next section. Additionally, 7.5% of the claims were categorized as partially verified. This category encompasses cases where insightful data were accompanied by incomplete interpretations, or where incomplete data were paired with insightful interpretations. Importantly, 6.8% of claims (69 out of 1006) were challenged. Among these, 44.9% (31 out of 69) were contested by published articles including 7.2% by the same authors, and 55.1% (38 out of 69) were challenged experimentally as part of the ReproSci project. Thus, the experimental validation performed during this project doubled the number of reported challenged claims. Finally, 1.2% of major claims were classified as “mixed” (Table 2). This category includes claims that presented particularly challenging issues (e.g., conflicting data), making it impossible to definitively categorize them as either verified or challenged.

Flow of the study.

Claims were first classified in five categories based on literature. 45 unchallenged claims were selected to be experimentally tested an classified either as challenged or as verified. We further classify some unchallenged as consistent or inconsistent depending on their consistency with current knowledge.s

In conclusion, contrary to the prevailing “reproducibility crisis” narrative, a significant proportion of the findings published in the field of Drosophila immunity before 2011 has been subsequently verified by independent studies.

A significant fraction of unchallenged claims is non-reproducible

Our analysis highlights a significant proportion of “unchallenged” claims that were not followed by subsequent studies and whose validity was unknown prior to our experimental evaluation. Most reproducibility studies have focused on verified or challenged claims so little is known about the vast number of statements generated in a scientific community without subsequent direct validation.

Despite the lack of validation, these “unchallenged claims” are often assumed valid and contribute to the broader construction of scientific knowledge. An important challenge is to estimate the percentage of unchallenged major claims that are actually non-reproducible.

Prior to our experimental validation, we identified a total of 285 unchallenged claims, representing 28.3% of all major claims published up to 2011. We conducted experimental evaluations of 45 unchallenged major claims with the assistance of multiple laboratories (Westlake et al., 2025). The selection of unchallenged claims for testing was intentionally biased toward two categories: (i) those that appeared suspicious in light of the existing literature (31 claims), and (ii) those that were relatively straightforward to test experimentally (14 claims). Among the 45 unchallenged major claims tested, 84% (38 out of 45) could not be experimentally reproduced and were then reclassified as “challenged by the ReproSci project.” Conversely, only 16% (7 out of 45) were successfully reproduced and reclassified as “verified by the ReproSci project” (Figure 1).

Notably, all 31 suspicious unchallenged claims were challenged, while 7 out of the 14 unchallenged “easy to test” claims were also challenged. Excluding the 45 unchallenged major claims that were experimentally tested, we categorized the remaining 240 unchallenged claims into three groups: 1) unchallenged logically consistent – claims not directly tested but consistent with current knowledge of the field (111 claims); 2) unchallenged logically inconsistent – claims that appear inconsistent with current knowledge (22 claims); 3) Unclassified – claims that could not be definitively categorized (107 claims) (Figure 1). In the remainder of the text, we analyze the claims following ReproSci experimental validation, labelling the 45 selected claims as either challenged or verified.

This analysis reveals that a substantial number of claims remain unchallenged even a decade after publication. Moreover, our findings suggest that a considerable proportion of these claims may not be reproducible.

Higher representation of challenged claims in trophy journals and from top universities

Articles published in high-impact journals receive high visibility, a key factor in career progression and funding security (van Wesel, 2016). Additionally, high-impact publications are increasingly used as metrics to evaluate university rankings. Consequently, scientists face significant internal and external pressures to publish in such prestigious journals (Udesky, 2025). On the other hand, high-impact journals aim to attract the most surprising findings, which are inherently more likely to be challenged than more conventional studies. Based on this, we hypothesized that articles published in high-impact journals are more likely to include non-reproducible results. To test this hypothesis, we estimated the number of challenged claims published in journals categorized as having low or medium-impact factors (<10), high-impact factors (>10 and <50), and “trophy journals” of the highest reputation (>50: Science, Nature, and Cell) in the Drosophila immunity at that time. The results are summarized in Figure 2, with a more detailed breakdown by journal provided in Table S1. Compared with low-impact journals, where 5.8 % of major claims were challenged, the proportion rose to 7.8 % in high-impact journals (odds ratio [OR] = 1.36, 95 % confidence interval [CI] 0.80–2.32) and more than doubled to 12.9 % in trophy journals (OR = 2.39, 95 % CI 1.06–5.40). We recall that the OR expresses how much more (OR > 1) or less (OR < 1) likely a claim is to be challenged relative to articles published in low-impact journals, while the 95 % CI indicates the range of values compatible with random sampling error; OR intervals that exclude 1 denote a statistically significant effect. Thus, our analysis show major claims published in low-impact journals are significantly more likely to be reproducible than major claims published in trophy journals. One explanation for this trend is that claims published in high-impact journals are more frequently subjected to replication attempts or used as a foundation for further research, thereby increasing the likelihood of identifying irreproducible results. Indeed, low-impact journals include a higher share of claims remaining unchallenged (29.3 %, 95 % CI 25.9–32.9 %) compared to high-impact journals (13.9 %, 95 % CI 10.5–18.2 %) and to trophy journals (17.7 %, 95 % CI 10.2–29.0 %). However, variation in the number of unchallenged claims does explain the higher proportion of challenged claims in trophy journals compared to high-impact journals. If we take into consideration the percentage of verified, unchallenged and challenged claims, Figure 2B shows that high-impact but not trophy or low-impact journals contain the highest proportion of verified claims. Thus, we conclude that while trophy journals may indeed attract some of the best science, they also appear more likely to include irreproducible claims.

Percentage of claims by category according to journal-impact and institution ranking.

(A, B) The distribution of claims across the five assessed categories for articles published in low-impact, high-impact, and trophy journals. The number of claims is indicated at the top. Panel A presents raw counts, while Panel B shows normalized data. (C) The distribution of claims across the five assessed categories according to the ranking of the last author institution in Shanghai Ranking’s 2010 Academic Ranking of World Universities. The number of claims is indicated at the top, while the number of institutions in each category is indicated at the bottom. The Not Ranked categories encompass research institutes (ex. CNRS), hospitals and other institution that are not included in the Shanghai ranking. In this figure, and in the rest of the paper, the number nc indicates the number of claims in each category, while the number at the bottom highlight the count in each category (e.g there are 11 high-impact journals, and 73 Top-50 institutions).

We then investigated whether the percentage of irreproducibility varies among laboratories affiliated according to the prestige of their affiliate universities. To test this hypothesis, we classified all the major claims published according to the Shanghai ranking of the last author university. Figure 2C shows that claims from laboratories at top-50 universities were challenged far more often 15.8 % (95 % CI 11.4–21.6 %) than other institutions, up to a four-fold increase in the odds of being questioned (OR = 4.69, 95 % CI 2.53–8.70) versus 3.9 % for unranked institutions, which notably include hospitals and research institutes such as the French CNRS. Even after adjusting for journal tier (see model SI1), university prestige remains the dominant signal: claims from top-50 institutions are 3.6 times more likely to be challenged than those from unranked universities (OR 3.57, 95 % HDI 1.67–7.92). By comparison, publishing in a “trophy” journal adds only a notable and uncertain risk (OR 1.63, 0.66–3.88), while other journal or ranking bands show no clear effect. Put simply, elite affiliation more than journal prestige best predicts irreproducibility in our sample.

The irreproducibility rate has increased over time as the field has grown in popularity

The reproducibility crisis narrative suggests that irreproducibility has increased over recent decades, posing a significant challenge to the life sciences (Baker, 2016; Hunter, 2017). Our dataset spans 50 years, from the onset of Drosophila immunity research as a relatively marginal field to its raise to prominence, with a notable increase in publications since the mid-1990s driven by growing interest in innate immunity. We analyzed fluctuations in irreproducibility over time by grouping major claim assessments into time windows: pre-1991, followed by 5-year intervals until 2011. The percentage of irreproducibility was calculated for each time window (Figure 3). Across the five windows examined, the proportion of challenged claims climbed steadily from 0% in the small corpus prior to 1991 (0/29 articles), to 4.6% during 1992–1996, and then further increased to 7.7% between 1997 and 2001. This upward trend peaked at 8.3% during the 2002–2006 period, before slightly declining to 6.1% between 2007 and 2011 (Figure 3). Despite these apparent temporal shifts, none of the pairwise comparisons between periods reached statistical significance, with p-values consistently ≥ 0.15. Thus, while the trends highlighted in our study are consistent with a raise of irreproducibility, we cannot exclude that the observed upward trajectory might reflect sampling variation rather than a definitive sustained temporal trend. Further studies focusing on articles published after 2011 could clarify whether the mild decrease we observed marks the beginning of a downward trend or reflects fluctuations at a consistently high level.

Evolution of claim replicability over time.

(A, B) The distribution of claims by categories of assessment over time is shown in A, and the proportion normalized over the total number of claims per time period is shown in B.

Interestingly, we observed a peak in verified claims between 1992 and 2001, accompanied by a lower percentage of unchallenged claims (Figure 3). This period coincided with the early expansion of the field when many researchers worked on similar research questions with a relatively narrow conceptual framework of innate immunity. The subsequent increase in unchallenged claims may reflect the rapid conceptual expansion of the field, which likely outpaced the growth in the number of researchers, reducing opportunities for overlapping or validating research efforts. An alternative explanation for this trend is that in the early stages of the field, researchers uncovered the largest, easily detectable effects. As the field matured, focus shifted to subtler phenotypes and complex regulatory mechanisms, making replication more challenging.

Overall, our time-course analysis suggests that irreproducibility has increased over the decades, aligning with the reproducibility crisis narrative. However, an alternative hypothesis is that irreproducibility varied alongside the field’s popularity, peaking once when the field became popular and had attracted newcomers who may not have been fully acquainted with its established standards or who joined the field by opportunistically. Furthermore, the rise in the percentage of challenged claims was accompanied by an increase in unchallenged claims, indicating a reduced degree of validation within the field.

First-author patterns of irreproducibility

Having analyzed the pattern of irreproducibility across a large number of articles, representing contributions from multiple laboratories over five decades, we next investigated the distribution of irreproducibility among authors and laboratories. A central question is whether irreproducibility is evenly distributed across the scientific community, or if hubs — specific authors or laboratories — are responsible for producing more invalid statements. First, we explored whether reproducibility is associated with the first author, who typically conducts the work, or with the last author, who manages the laboratory and may foster a research environment that produces irreproducible results.

To explore the relationship between irreproducibility and first authorship, we computed the percentage of “challenged” articles attributed to each first author. Additional information on first author was obtained through interviews with principal investigators (PIs) or via public LinkedIn profiles, including variables such as gender and position in the laboratory.

In total, 289 first authors contributed to the 400 articles analyzed by the ReproSci project, with an average of 3.45 claims per author. Notably, 47 authors (16.3%, 95% CI: 11.4–19.6%) published at least one non-reproducible claim, with one author reaching up to 83% irreproducibility (5 challenged claims out of 6 from two articles). Figure 4A displays the distribution of challenged claims among first authors. A statistical dispersion analysis revealed a Gini index of 0.88 (inequality index ranging from 0 -perfect equality- to 1 -extreme inequality-), where the top 10% of first authors accounted for 73.9% of all challenged claims, and the top 20% accounted for all 100% of challenged claims (Figure 4B), suggesting a highly unequal distribution of irreproducibility according to first author.

Distribution of irreproducibility according to first authors.

A) Distribution of irreproducible claims among first author. We note that all irreproducible claims are produced by less than 20% of the first authors. B) Distribution of irreproducibility for all first authors. Each dot represents an author, ranked according to the proportion of challenged claims. The size of the circle indicates the number of claims and the color the percentage of verified claims.

We next investigated whether irreproducibility varied by the gender of the first author. The proportions of challenged articles were similar for male and female first authors: 7.3% and 6.7%, respectively, and this difference was not statistically significant (OR = 1.09, 95% CI: 0.66–1.78, Fisher’s exact p = 0.802; Figure 5A).

Influence of parameters associated with first author on reproducibility.

A) Repartition of claims in each assessment category according to the gender of the first authors. The gender of four authors could not be inferred based on first name ambiguity explaining why the total number of authors is 285. B) Repartition of claims in each assessment category according to status of the first authors. The stages of career of 279 first authors were obtained by interviewing last authors, querying Linkedln public profiles or extrapolated according to the publication list. Senior staff refers to permanent position scientist working in a lab. ND stands for Not Determined. C) Proportion of challenged claims as a function of the number of published articles for first-authors. The color indicates the percentage of claims that are verified.

We were able to determine the status of most of first authors at the time of the article was published, most being PhD students (n=168) or postdoctoral researchers (n=87), with smaller numbers identified as technicians (n=5), Masters students (n=5), PIs (n=8), or permanent staff (n=6) (Figure 5B). No major differences in irreproducibility were observed between PhD students and post-docs. While we did observe a higher proportion of challenged claims among PIs and Master’s students, and a lower proportion among technicians and permanent staff, these trends are difficult to interpret due to small sample sizes.

Finally, we tested the relation between the number of articles published by a first author and the proportion of irreproducible research. No significant trend was observed (linear regression; r2=0.01, p=0.088; Figure 5C), suggesting the level of author productivity is independent of irreproducibility. In summary, we observed considerable variability in irreproducibility among first authors, suggesting that individual authors may substantially contribute to reproducibility issues.

Lead authors pattern of irreproducibility

We now turn to the question of whether irreproducibility varies according to laboratory. As it is common practice in biology, we assigned the last author (referred to as PI for principal investigator) to represent the laboratory. The 400 articles analyzed were co-authored by 156 leading authors from various institutions, predominantly from France, Sweden, Japan, South Korea, and the USA. For each last author, we estimated the number of major claims in each of the five categories. The variation in irreproducibility across these PIs is shown in Figure 6A. A statistical dispersion analysis in Figure 6B revealed a Gini coefficient of 0.856: the top 10% of all last authors accounted for 71.0% of all challenged claims, while the top 20% accounted for 94.2%, suggesting a highly unequal distribution of irreproducibility according to laboratory. We next explored whether there were differences in irreproducibility between male and female last authors. The percentage of challenged claims was slightly higher for male last authors (7.0%, 95% CI: 7.0%–7.1%) compared to female last authors (5.8%, 95% CI: 5.7%–6.0%), but this difference was not statistically significant (OR = 1.22, 95% CI: 0.59–2.51; Fisher’s exact p = 0.73; Figure 7A).

Distribution of irreproducibility according to last authors (PI).

A) Distribution of irreproducible claims among leading authors. B) Distribution of irreproducibility for all leading authors. Each dot represents a leading author, ranked according to the proportion of challenged claims. The size of the circle indicates the number of claims and the color the percentage of verified claims.

Influence of various parameters associated with the last author on reproducibility.

A) Distribution of claims by category according to the gender of the last authors. B) Distribution of claims by category according to the seniority of the last authors. Senior PI were defined as having published a last author article at least 5 years before the considered articles. The total of junior and senior authors is higher than the total of last authors as some PIs published both as junior and senior. C) Proportion of challenged claims by last author as a function of number of articles.

Pressure to publish is typically stronger among junior PIs who often lack tenure and must establish their laboratory. To examine whether this pressure might be associated with irreproducibility, we classified PIs as junior or senior based on whether they had published a last-author article at least five years prior to the considered publication. Although Figure 7B indicates a slightly higher proportion of challenged claims among junior PIs (7.2%, 95% CI: 7.2%–7.3%) compared to senior PIs (6.5%, 95% CI: 6.5%–6.6%), this difference was not statistically significant (OR = 1.12, 95% CI: 0.68–1.83, Fisher’s exact p = 0.70). This modest difference was further attenuated when considering mixed and partially verified claims.

We then explored whether the percentage of irreproducibility tended to decline according to the number of published articles as last author (Figure 7C). We could hypothesize that PIs that have transiently work on Drosophila immunity would have more irreproducible claims, or on the contrary that the most productive laboratories compromise speed over rigor. However, no significant trend was observed between the percentage of challenged claims and the number of articles published by an author (linear regression; r2=0.004, p=0.67).

We found that the percentage of irreproducibility has increased over time as the field became more popular. We therefore explored if this increase in irreproducibility affects all PIs or only new PIs that joined the field at a later time. We show in Figure 8A the proportion of challenged claims by last authors according to the publication date of their first article on Drosophila immunity, either as last or as first authors. We observed a significant association between irreproducibility and the period when principal investigators (PIs) established their laboratories. Specifically, PIs, who started their laboratories after 1995, had a significantly higher percentage of challenged claims (8.1%, 95% CI: 8.0%–8.1%) compared to those who began their labs before 1995 (2.6%, 95% CI: 2.6%–2.7%; OR = 3.24, 95% CI: 1.38–7.59, Fisher’s exact p < 0.001).

Patterns of irreproducibility by last author according to time of starting their laboratory and research style.

A. The proportion of challenged claims by last author is shown according to the time at which the author published their first paper as a first or last authors. The increase in irreproducibility is associated with PIs who started their lab after 1995. Circle size indicates the number of published articles by each author. B. Distribution of claims by last author according to whether they have published at least one first author article on Drosophila immunity in another laboratory. C. Distribution of claims according to research style. Proportion of challenged claims by PIs who have continuously work on the Drosophila immunity (‘continuity PIs’) or made a transient incursion into the field (‘exploratory PIs’). Only PIs, who published as last authors after 1995, were taken into consideration.

Among the PIs working in the Drosophila immunity field, who established their lab after 1995, we identified 22 who previously published at least one first-author article in this field during their PhD or postdoctoral training in another laboratory. We hypothesized that these authors, having gained prior hands-on experience on Drosophila immunity, would be less prone to publishing irreproducible claims. Consistent with this hypothesis, we observed significantly lower levels of irreproducibility among these experienced authors (3.5%, 95% CI: 3.4%–3.5%) compared to authors who had not been trained in laboratories working on Drosophila immunity (8.2%, 95% CI: 8.1%– 8.2%; OR = 2.48, 95% CI: 1.19–5.18, Fisher’s exact p = 0.01; Figure 8B).

We conclude that irreproducibility varies significantly across laboratories with a limited number of laboratories accounting for most of the irreproducible claims. As for first authors, author gender is not significantly associated with reproducibility. Our studies point to a significant role of experience: PIs that had previously worked in another Drosophila immunity laboratory to the extent that they published a first author article in that laboratory were less prone to publish irreproducible claims.

Irreproducibility according to research styles

Among the various career progression strategies, one involves deepening expertise and specialization within the same field, while another entails transitioning to different fields or exploring different career paths. This may reflect the personality of the scientist, either opting for a more detailed, focused approach or seeking continual new challenges. Scientists from the past were often spending their life on a specific topic within a field. In the last decades, changing field has become the norm (Zeng et al., 2019). Accordingly, scientists are not funded by institutions to work on a topic long-term anymore, but rather funding is granted by agencies for specific short-term projects. Project grants define research with the primary goal to do ‘breakthrough discoveries’, not accumulative research, which is sometimes treated as the ‘pedestrian’ or the ‘stamp collector’ way of doing research. This leads scientists to move quickly into new research fields in search for opportunities of funding and recognition, a nuanced hypothesis supported by the recent description of a “pivot penalty” for scientists that change career tracks (Hill et al., 2025). We can define these two strategies as either “continuity PIs,” which spend their career on one core topic, and “exploratory PIs”, which regularly venture into new topics. To further explore this question, we classified all last authors that set up their laboratory after 1995 as ‘continuity’ PIs or ‘exploratory’ PIs. Exploratory PIs are defined as those who made a transient incursion in the Drosophila immunity field before moving on to another field or leaving academia. In contrast continuity PIs worked continuously in the Drosophila immunity field throughout their careers. We observed that continuity PIs had a lower proportion of challenged claims (6.1%, 95% CI: 6.0%–6.2%) compared to exploratory PIs (8.2%, 95% CI: 8.1%–8.3%). However, this difference was not statistically significant (OR = 1.38, 95% CI: 0.84–2.26, Fisher’s exact p = 0.24; Figure 8C). We also noted a markedly higher share of unchallenged claims among exploratory PIs than among continuity PIs, 32.2 % (95 % CI 32.1– 32.4 %) versus 19.1 % (95 % CI 19.0–19.2 %), respectively. This difference was statistically significant (OR = 2.02, 95 % CI 1.50–2.71; Fisher’s exact p < 0.001), reinforcing the idea that researchers, who move on to new topics, tend to leave a larger fraction of their earlier findings without follow-up or validation by the community. Since unchallenged claims are more likely to be irreproducible, this suggests that irreproducibility is higher in the work of PIs with an exploratory style, as opposed to those engaged in a continuity-driven research style.

Multivariable analysis of predictors of claim irreproducibility

We have presented above a comprehensive exploratory data analysis, examining one variable at a time. However, drawing robust conclusions necessitates the joint analysis of all covariates, while appropriately accounting for the non-independence of claims made by the same authors. Thus, we fitted a multivariate Bayesian mixed-effects logistic model to uncover the effect of each covariate, while taking into account all other confounding variables at our disposal (Abril-Pla et al., 2023; Capretto et al., 2022)(Figure 10). The model simultaneously adjusts for author demographics, laboratory attributes and journal or institutional prestige and predict a binary outcome for each claim: challenged or not challenged (includings Verified, Partially Verified, Mixed and Unchallenged) (See in Supplementary figure S3 a similar model to predict unchallenged or not unchallenged claims). As the claims from the same first or last author are not independent, we included a mixed effect for each first author (289) and each last author (156). These effects capture the influence of a particular author in driving a claim reproducibility. We conducted extensive test on model formulation and convergence (see methods, and Supplementary information for posterior predictive and diagnostics plot). Note that as this analysis is Bayesian, we report the Highest Density Interval (HDI) instead of the confidence interval. We observe that, none of the first-author factors (gender: OR= 0.94, 94% HDI: 0.42–2.16; trainee stage: OR = 1.17, 94% HDI: 0.51–2.87) or leading-author factors (gender, seniority, continuity, or having entered the field before 1995) showed a decisive association with irreproducibility once every other covariate was taken into account (all HDIs span 1). We still observe a trend toward reduction of challenged claims for lead authors having being trained in Drosophila immunity or have started their laboratory before 1995. Figure S3 shows that lead author with an exploratory life style tends to have significantly more unchallenged claims (OR: 2.22, 1.18, 4.55). In contrast, university ranking and journal level covariates dominated the model: articles from top-50 universities were almost four times more likely to contain a challenged claim than those from un-ranked institutions (OR 3.72, 1.55–9.43), and papers in trophy journals (Science, Nature, Cell) carried an elevated uncertain risk (OR 1.76, 0.62–5.00) compared to low-impact journals.

Multivariable analysis of predictors of claim irreproducibility.

Intermediate effects for high-impact journals and mid-ranked universities were positive but imprecise. Publication year was modelled with a three-knot B-spline to allow for non-linear time trends. None of the individual spline coefficients differed credibly from zero, indicating that once journal tier and the other covariates were included the residual calendar-year effect on irreproducibility was negligible, indicating that time of publication itself is not a major factor.

The hierarchical component comprised 289 random intercepts for first authors and 156 for PIs. The posterior distribution of these intercepts (see Supplementary Information) is broad, confirming that some individuals or laboratories have a systematically higher or lower propensity for challenged claims. Nevertheless, only the three most extreme first-author effects had 95 % highest-density intervals that excluded zero (and all with positive irreproducibility) meaning that statistically robust clustering of irreproducibility is confined to a handful of contributors; for the vast majority of authors the collected evidence does not support significant differences in reproducibility from what is expected from a random distribution.

Taken together, the model confirms our exploratory data analysis, namely that it is the prestige of institution and publication in trophy journals rather than the personal characteristics of the scientists that best predicts which claims will later prove to be irreproducible.

The forest plot reports the odds ratio and the 94% highest density interval of each covariate effect on a claim becoming challenged.

Discussion

The ReproSci project presents an in-depth analysis of reproducibility in an experimental science field with a 14-year retrospective period. This initiative focuses specifically on conceptual replicability —the ability to independently reproduce key findings based on the information provided— rather than direct reproducibility using the exact same data and conditions (Plesser, 2018). Replicability is of critical importance to the scientific community who seeks to build upon published claims. One of our project’s key strengths lies in its comprehensive scope: it spans over fifty years of research, from 1959 to 2011, within a specific scientific community, resulting in a robust database that synthesizes the main findings across decades. This time frame encompasses the formative years of the field, characterized by a small number of pioneering laboratories, followed by a period of rapid expansion driven by increased interest, heightened competition, and publication in high-impact journals. This growth phase was marked by a proliferation of knowledge and conceptual innovation, with novel approaches to studying the immune system emerging. With the awarding of the Nobel Prize in 2011 and subsequently the shift toward more translational research, the growth pace of the Drosophila immunity field has since stabilized, and the field remains highly active.

Although our study centers on a specific experimental domain at a particular moment in time with a specific geographical development, it reveals broader patterns of irreproducibility that are likely generalizable. These findings contribute meaningfully to the ongoing conversation about the reproducibility crisis in the experimental life sciences (Fanelli, 2018; França and Monserrat, 2018; Hunter, 2017; Peng, 2015).

Contrary to the more dramatic narratives surrounding the reproducibility crisis (Amaral et al., 2025; Baker, 2015; Errington et al., 2021), our study shows that a substantial proportion of claims—61%— can be considered as verified. The proportion of claims that have been directly challenged remains relatively low (7%), though likely underestimated. This indicates that research in the Drosophila immunity field is rather solid, possibly due to the power of Drosophila genetics, the collective ethos of sharing within the community, a less competitive environment, and a lack of influence associated with translational research.

A novel and important contribution of our work is the identification of a large fraction of unchallenged claims —findings that, despite being published, have never been independently tested. The presence of these unchallenged claims could result from the rapid expansion of the field limiting overlapping research that would validate them. However, the existence of unchallenged claims could also be explained if they are not reproducible. For instance, the absence of follow-up studies could reflect authors that are active in the field being aware of the fragility of specific unchallenged claims, and so not pursuing them. The existence of unchallenged claims could also stem from the reluctance of other researchers to publish contradictory statement due to the stigma of publishing negative results, the so-called “file drawer” effect (Rosenthal, 1979). Indeed, many scientists report abandoning verification efforts when unable to reproduce a colleague’s findings, fearing personal conflict or professional retaliation (DeCoursey, 2006; Kannan and Gowri, 2014; Nissen et al., 2016). Our experimental validation demonstrates that a significant proportion of unchallenged claims are, in fact, non-reproducible. However, we should note that we did not randomly select the unchallenged claims that were to be experimentally tested, and we favored suspicious claims and claims that were easy to experimentally test. Thus, the true rate of reproducibility of unchallenged claims should not be directly extrapolated from our experimental validation attempts. However, we found that none of unchallenged, suspicious claims could be reproduced. While confirmation bias driven by our initial suspicion cannot be entirely excluded, the absence of follow-up in the published literature for over a decade later supports our interpretation that these findings were not robust. Of note, we considered many of these unchallenged claims were considered suspicious based on informal exchanges within research communities, that reported difficulties in reproducing the findings. Most of these non-reproducible suspicious claims were published in high-impact or trophy journals (26/31) with great visibility, perhaps contributing to this privately-shared stigma within the community. Taken together, our findings support the notion that “big stories” published without subsequent follow-up may be indicative of irreproducibility.

Another takeaway from our experimental validation is that approximately 50% of the unchallenged claims were not previously flagged as problematic failed the experimental validation test regardless. While a claim could remain unchallenged for years because it is valid and no one has anything novel to add to it, our results suggest the status of a claim being “unchallenged”is not a reliable proxy for its validity. If we reclassify the dataset by counting logically consistent unchallenged claims as “verified” and logically inconsistent ones as challenged, and conservatively assume that 50% of the remaining unchallenged claims fall into the latter, we estimate that the true proportion of challenged claims may approach 14%. When we include mixed and partially verified claims, a rough upper-bound estimate suggests that at least 80% of all published claims on Drosophila immunity between 1959-2011 are verified or likely to be true leaving 20% of claims either challenged or difficult to interpret. We will take this moment to remind that challenged claims can include work that was done well, but where the field has advanced in its understanding and can better interpret the original findings.We believe the percentage of challenged claims likely varies across scientific fields, depending on intrinsic challenges such as the complexity of biological processes, methodological variability, and the field’s attractiveness, which may increase competitive pressure.

In a recent editorial, Fiala and Diamendis, suggested that reproducibility efforts should focus on high-impact papers as these articles are most deleterious to the scientific community (Fiala and Diamandis, 2018). In our database, we indeed found that “trophy journals”, considered to contain the most impactful science, are more likely to publish articles with challenged claims than high or low-impact, although the difference was not significant. Three intertwined factors can explain why high-impact factor journals have higher rate of non-reproducibility. First, the uncertainty that comes with exploring the frontiers of knowledge may lead to investigations that are founded on an incomplete understanding of the major factors influencing the results. Second, the unconventional nature of some high-impact discoveries, and the fact that claims reported in these journals generally attract more attention, increases the likelihood that other researchers will attempt to replicate them — thereby raising the number of challenged claims. And third, systemic biases that pressure scientists to publish in these journals for prestige might encourage scientists to take shortcuts in finalizing their studies, increasing the chance of challenged claims. In our study, we observed a decreased proportion of unchallenged claims in articles published in high-impact and trophy journals indicating that claims published in these journals are more reproduced. However, this higher rate of follow-up work cannot fully explain by itself the higher proportion of challenged claims in trophy journals.

Compared to low rank and trophy journals, high-ranking journals have the best ratio of verified to challenged claims, as low-impact journals have a high percentage of unchallenged claims. Strikingly, we observed higher rate of non-reproducible claims for scientists working in universities in the top 50 Shanghai ranking. Previous work has suggested journal rank as a major associate of irreproducibility (Brembs et al., 2013), and that high-impact journals favour authors from leading institutions (Kulal et al., 2025). Our study highlights how university ranking is among the most striking correlates of irreproducibility. Taken together, the cause and effect relationship at play here may be driven most by factors associated with institutional prestige, such as trust among networks of colleagues that act as editors and reviewers, leading to greater acceptance rates at high-impact journals. Importantly, our results suggest this greater acceptance does not strictly reflect a high quality of the work being published, but rather a willingness to trust irreproducible work more frequently if it comes from a prestigious institute. Our study spans a period during which most research was done in North America, Europe, Japan and South Korea and does not capture the rise of science in China, or rapid growth of scientific output across the global north and south, where scientists in some countries face strong incentive for ‘high impact’ publications (Delgado-López-Cózar et al., 2021; Quan et al., 2017). The fact that the Drosophila immunity field monitored in this study developed initially in Sweden and France, which have unique geographic and research community characteristics, might affect some our conclusions or limit them to a specific period.

Future studies in other areas of research might help to confirm if working in top ranked universities correlates with higher rate of non-reproducible science.

Our data align with the broader narrative of increasing rates of non-reproducible science. One possible explanation is a general decline in the standards of rigor in recent years in the life sciences; however, a non-mutually exclusive hypothesis is that the rise in irreproducibility may reflect the increasing popularity of certain fields, creating greater variance by attracting more researchers, who themselves may not be fully acquainted with the field’s established norms and methodological rigor, or who move opportunistically from field to field to publish striking papers. When research on the gut microbiota gained momentum in the 1990s, it was rapidly positioned as a promising avenue to explain a wide array of diseases. Substantial funding and media attention enabled the swift development of the field. However, many early studies lacked methodological rigor and could not be reproduced (Wesolowska-Andersen et al., 2014). The lower scientific standards tolerated in this emerging area —compared to more established fields— facilitated the publication of high-profile articles, encouraging a rapid influx of new research groups. In this context, scientific novelty and perceived importance often took precedence over precision and reproducibility. A similar but more acute pattern was observed during the SARS-CoV-2 pandemic. Alternatively, this reproducibility crisis might be more linked to specific features of our time. Although this notion is disputed (Lemaitre, 2016; Twenge and Campbell, 2009), an increase of narcissism in society might increase the chance of distorting science to favor our own success.

Importantly, our analysis shows that irreproducibility is not evenly distributed across all researchers. While irreproducibility can stem from the inherent difficulty or novelty of certain research questions, our findings point to the significant influence of both the first author (who performs the work) and the laboratory environment. Surprisingly, we did not observe a major effect of experience level: irreproducibility rates were similar between PhD students and postdoctoral researchers and did not markedly decrease with the number of publications. Similarly, we only observe a slightly but not significant higher irreproducibility rate among junior faculty compared to senior professors. This suggests that irreproducibility is not inherent to early career stage but rather to certain research practices and environments. That said, we did observe a higher irreproducibility rate among PIs that entered the field from a non-Drosophila immunity background. This observation supports the value of traditional scientific training pathways, where early-career researchers are mentored in well-established environments before gaining independence.

Another noteworthy finding is our suggested link between irreproducibility and “continuous” or “exploratory” research style. We observed that authors with a more exploratory research style have slightly higher proportions of challenged claims and a significantly higher proportion of unchallenged claims, the latter being more likely to be challenged as shown by our experimental replication. Exploratory researchers have worked only transiently in the field, often leaving it after its peak period of popularity. In recent years funding agencies and universities favored this exploratory research style focusing on breakthroughs, notably by funding projects that may be rapidly discontinued. As a counterpoint of this, our study highlights the value and reliability of laboratories that tend to maintain long-term commitment to a specific field.

An important aspiration in the field of metascience study would be the ability to predict claims or articles that are likely irreproducible. Our retrospective analysis identifies several criteria that correlate with irreproducibility, although not all of them reach significance. These include i) articles being published by authors working in top-ranked institutions, ii) articles being published in trophy journals, iii) the absence of follow-up studies (i.e. being ‘unchallenged’), and iv) articles being published by authors with an exploratory research style.In contrast, having published as first author in another laboratory within the field, and publishing in high-impact journals but not trophy journals correlate with higher reproducibility. The absence of follow-up studies regarding claims published in high-impact journals is likely the most useful proxy to question the reproducibility of these claims.

Limitations

This work has several limitations. First, the definition of the claim, the validity, the determination of the validation was influenced by the annotating authors (HW and BL). While the ReproSci website was opened to the community, many PIs were reluctant or not interested in the project of testing old claims that represent “cold scientific cases”. Second the leading author is stakeholder of the project having published several articles, that were included in the analysis, including 44 articles with B.L. as co-author with 4 as first and 22 as last authors. Thus, this may not represent the general view of the community. Finally, the findings might be specific to this field and strongly influenced by specific factors idiosyncratic to its development. However, we believe that despite these limitations the systematic and retrospective analysis of a research field provides insights to better apprehend how science progresses despite errors and the impact of irreproducibility.

Methods

Selected articles, annotation, experimental validation of the ReproSci project

We provide here a succinct method of the ReproSci reproducibility project (see the companion article (Westlake et al., 2025) for more details). In brief, a list of 400 publications published before 2011 was generated using a curated search string on the publicly available PubMed database. This limitation was imposed to allow a time frame in which follow-up studies may be published that may verify, expand on or contradict the content of the publications selected for the study. Selected primary articles were annotated by a single researcher (HW) and reviewed by a second researcher (BL) to ensure consistency and accuracy across the project. An online database, ReproSci (https://ReproSci.epfl.ch/), was constructed to organize the annotations and provide an interactive and searchable interface. Claims were extracted through careful reading of the selected primary articles and organized according to their importance within the paper: a single Main claim representing the title claim of the paper or overarching finding; 3-4 Major claims emphasized in the abstract or those contributing to the overarching finding; and a variable number of Minor claims representing other findings of note not emphasized in the abstract. Claims were cross-checked with evidence from previous, contemporary and subsequent publications and assigned a verification category (see Table 1). A short statement explaining the evidence and citing sources was written to justify the categorization and provide additional context for the claim; this assessment was linked to each claim and can be viewed on the ReproSci website. A subset of unchallenged claims (45 major and 11 minor) was selected for direct experimental reproduction by various authors from different laboratories with relevant expertise. This project is concerned with the accuracy of the conclusions themselves rather than the reliability of methods, so we employed indirect or inferential reproducibility, where experimental procedures that may be different from those originally employed are used to independently verify the claim. The ReproSci website was made available to the Drosophila scientific community on July 11, 2023, with an email to encourage to comment on the annotations and verifications and contribute evidence.

First authors and last author classification

Information regarding the gender and status of first authors were either obtained by interviewing PIs (for 191 first authors) or found on publicly available websites (LinkedIn and university websites) or were implied based on their first name usage (gender) and their PubMed publication lists (for the other 100 first authors). When some of these information could not be retrieved, these authors were not taken in consideration.

PIs were manually classified as senior or junior whether they have or have not published a last author article at least five years before the considered publication. Continuity and exploratory PIs were manually classified according to their PubMed publication list. Continuity PIs are PIs that have continued to work in the Drosophila immunity field until now or to the end of their scientific career (deceased, retired, left academia). Exploratory PIs are PIs that have done a transient incursion in the Drosophila immunity field moving to other topics of research.

Statistical analysis

The analysis was done on the 1006 major claims of the 400 articles. The cleaned database contains one row per major claim with linked article-, author- and laboratory-level covariates, which was subsequently analyzed. Every output, counts, proportions and graphics, exploratory summaries and Bayesian mixed-effects models are defined under a reproducible Python workflow, which enable the full replication of the main results and every figure in the manuscript.

The ninety-five-percent confidence intervals (CI) for single proportions were determined using Wilson’s score method. Whenever a binary predictor (e.g. author gender) was cross-classified with the binary outcome “claim challenged”, we reported the odds ratio with its 95% CI and the two-sided p-value from Fisher’s exact test; The raw contingency tables for every binary predictor appear in Supplementary Document.

Multivariable modelling

Two Bayesian regressions were fitted with weakly-informative priors. Posterior summaries are presented as odds ratios (OR) with 94 % highest-density intervals (HDI).

Hierarchical logistic regression for claim irreproducibility. The outcome was whether a major claim was later classified as challenged. Fixed-effect predictors captured article prestige, author characteristics and laboratory context: journal tier (Low, High, Trophy), university-ranking tier (Not ranked, 101 +, 51–100, Top 50), publication year modelled with a three-knot B-spline, first-author gender and training stage (PhD vs post-doc), last-author gender, laboratory seniority (junior vs senior PI), continuity style (continuous vs exploratory) and whether the laboratory had published in the field before 1995. Claims from one author not independent, so random intercepts were included for both first-author key and last-author key.

Another simpler hierarchical regression model was used with only authors random intercepts, university ranking and journal impact factor to help the conclusion in section Supplementary Figure S2.

All models were estimated with the dynamic-HMC NUTS sampler (target _accept = 0.9); four independent chains of 2 000 post-warm-up draws were run for every model. Convergence was confirmed both visually and numerically (all Rhat≤1.01; all effective sample sizes > 400).

Supplementary Fig. S4 shows trace plots, posterior-predictive checks and Pareto-k diagnostics; no systematic lack of fit was detected. Exact formulas, priors, sampling statements and notebooks are available in the public repository (Supplementary Document).

eLife Assessment

This important study presents an impressive large-scale effort to assess the reproducibility of published findings in the field of Drosophila immunity. The authors analyse 400 papers published between 1959 and 2011, and assess how many of the claims in these papers have been tested in subsequent publications. In a companion article they report the results of experiments to test a subset of the claims that, according to the literature, have not been tested. The present article also explores if various factors related to authors, institutions and journals influence reproducibility in this field. The evidence supporting the claims is solid, but there is considerable scope for strengthening and extending the analysis. The limitations inherent to evaluating reproducibility based on the published literature should also be acknowledged.

Reviewer #1 (Public review):

Summary:

The authors set out on the ambitious task of establishing the reproducibility of claims from the Drosophila immunity literature. Starting out from a corpus of 400 articles from 1959 and 2011, the authors sought to determine whether their claims were confirmed or contradicted by previous or subsequent publications. Additionally, they actively sought to replicate a subset of the claims for which no previous replications were available (although this set was not representative of the whole sample, as the authors focused on suspicious and/or easily testable claims). The focus of the article is on inferential reproducibility; thus, methods don't necessarily map exactly to the original ones.

The authors present a large-scale analysis of the individual replication findings, which are presented in a companion article (Westlake et al., 2025. DOI 10.1101/2025.07.07.663442). In their retrospective analysis of reproducibility, the authors find that 61% of the original claims were verified by the literature, 7.5% were partialy verified, and only 6.8% were challenged, with 23.8% having no replication available. This is in stark contrast with the result of their prospective replications, in which only 16% of claims were successfully reproduced.

The authors proceed to investigate correlates of replicability, with the most consistent finding being that findings stemming from higher-ranked universities (and possibly from very high impact journals) were more likely to be challenged.

Strengths:

(1) The work presents a large-scale, in-depth analysis of a particular field of science that includes authors with deep domain expertise of the field. This is a rare endeavour to establish the reproducibility of a particular subfield of science, and I'd argue that we need many more of these in different areas.

(2) The project was built on a collaborative basis (https://ReproSci.epfl.ch/), using an online database (https://ReproSci.epfl.ch/), which was used to organize the annotations and comments of the community about the claims. The website remains online and can be a valuable resource to the Drosophila immunity community.

(3) Data and code are shared in the authors' GitHub repository, with a Jupyter notebook available to reproduce the results.

Main concerns:

(1) Although the authors claim that "Drosophila immunity claims are mostly replicable", this conclusion is strictly based on the retrospective analysis - in which around 84% of the claims for which a published verification attempt was found. This is in very stark contrast with the findings that the authors replicate prospectively, of which only 16% are verified.

Although this large discrepancy may be explained by the fact that the authors focused on unchallenged and suspicious claims (which seems to be their preferred explanation), an alternative hypothesis is that there is a large amount of confirmation bias in the Drosophila immunity literature, either because attempts to replicate previous findings tend to reach similar results due to researcher bias, or because results that validate previous findings are more likely to be published.

Both explanations are plausible (and, not being an expert in the field, I'd have a hard time estimating their relative probability), and in the absence of prospective replication of a systematic sample of claims - which could determine whether the replication rate for a random sample of claims is as high as that observed in the literature -, both should be considered in the manuscript.

(2) The fact that the analysis of factors correlating with reproducibility includes both prospective and retrospective replications also leads to the possibility of confusion bias in this analysis. If most of the challenged claims come from the authors' prospective replications, while most of the verified ones come from those that were replicated by the literature, it becomes unclear whether the identified factors are correlated with actual reproducibility of the claims or with the likelihood that a given claim will be tested by other authors and that this replication will be published.

(3) The methods are very brief for a project of this size, and many of the aspects in determining whether claims were conceptually replicated and how replications were set up are missing.

Some of these - such as the PubMed search string for the publications and a better description of the annotation process - are described in the companion article, but this could be more explicitly stated. Others, however, remain obscure. Statements such as "Claims were cross-checked with evidence from previous, contemporary and subsequent publications and assigned a verification category" summarize a very complex process for which more detail should be given - in particular because what constitutes inferential reproducibility is not a self-evident concept. And although I appreciate that what constitutes a replication is ultimately a case-by-case decision, a general description of the guidelines used by the authors to determine this should be provided. As these processes were done by one author and reviewed by another, it would also be useful to know the agreement rates between them to have a general sense of how reproducible the annotation process might be.

The same gap in methods descriptions holds for the prospective replications. How were labs selected, how were experimental protocols developed, and how was the validity of the experiments as a conceptual replication assessed? I understand that providing the methods for each individual replication is beyond the scope of the article, but a general description of how they were developed would be important.

(4) As far as I could tell, the large-scale analysis of the replication results was not preregistered, and many decisions seem somewhat ad hoc. In particular, the categorization of journals (e.g. low impact, high impact, "trophy") and universities (e.g. top 50, 51-100, 101+) relies on arbitrary thresholds, and it is unclear how much the results are dependent on these decisions, as no sensitivity analyses are provided.

Particularly, for analyses that correlate reproducibility with continuous variable (such as year of publication, impact factor or university ranking, I'd strongly favor using these variables as continuous variables in the analysis (e.g. using logistic regression) rather than performing pairwise comparisons between categories determined by arbitrary cutoffs. This would not only reduce the impact of arbitrary thresholds in the analysis, but would also increase statistical power in the univariate analyses (as the whole sample can be used in at once) and reduce the number of parameters in the multivariate model (as they will be included as a single variable rather than multiple dummy variables when there are more than two categories).

(5) The multivariate model used to investigate predictors of replicability includes unchallenged claims along with verified ones in the outcome, which seems like an odd decision. If the intention is to analyze which factors are correlated with reproducibility, it would make more sense to remove the unchallenged findings, as these are likely uninformative in this sense. In fact, based on the authors' own replications of unchallenged findings, they may be more likely to belong the "challenged" category than to the "unchallenged" one if they were to be verified.

Reviewer #2 (Public review):

Summary:

Lemaitre et al. conducted an analysis of 400 publications in the Drosophila immunity field (1959-2011), performing both univariable and multivariable analyses to identify factors that correlate with or influence the irreproducibility of scientific claims. Some of the findings are unexpected, for instance, neither the career stage of the PI nor that of the first author appears to matter that much, while others, such as the influence of institutional prestige or publication in "trophy journals," are more predictable. The results provide valuable insight into patterns of irreproducibility in academia and may help inform policies to improve research reproducibility in the field.

Strengths:

This study is based on a large, manually curated dataset, complemented by a companion paper (Westlake et al., 2025. DOI 10.1101/2025.07.07.663442) that provides additional details on experimentally documented cases. The statistical methods are appropriate, and the findings are both important and informative. The results are clearly presented and supported by accessible documentation through the ReproSci project.

Weaknesses:

The analysis is limited to a specific field (immunity) and model system (Drosophila). Since biological context may influence reproducibility -- for example, depending on whether mechanisms are more hardwired or variable -- and the model system itself may contribute to these effects (as the authors note), it remains unclear to what extent these findings generalize to other fields or organisms. The authors could expand the discussion to address the potential scope and limitations of the study's generalizability.

Reviewer #3 (Public review):

Summary:

The authors of this paper were trying to identify how reproducible, or not, their subfield (Drosophilia immunity) was since its inception over 50 years ago. This required identifying not only the papers, but the specific claims made in the paper, assessing if these claims were followed up in the literature, and if so whether the subsequent papers supported or refuted the original claim. In addition to this large manually curated effort, the authors further investigated some claims that were left unchallenged in the literature by conducting replications themselves. This provided a rich corpus of the subfield that could be investigated into what characteristics influence reproducibility.

Strengths:

A major strength of this study is the focus on a subfield, the detailing of identifying the main, major, and minor claims - which is a very challenging manual task - and then cataloging not only their assessment of if these claims were followed up in the literature, but also what characteristics might be contributing to reproducibility, which also included more manual effort to supplement the data that they were able to extract from the published papers. While this provides a rich dataset for analysis, there is a major weakness with this approach, which is not unique to this study.

Weaknesses:

The main weakness is relying heavily on the published literature as the source for if a claim was determined to be verified or not. There are many documented issues with this stemming from every field of research - such as publication bias, selective reporting, all the way to fraud. It's understandable why the authors took this approach - it is the only way to get at a breadth of the literature - however the flaw with this approach is it takes the literature as a solid ground truth, which it is not. At the same time, it is not reasonable to expect the authors to have conducted independent replications for all of the 400 papers they identified. However, there is a big difference trying to assess the reproducibility of the literature by using the literature as the 'ground truth' vs doing this independently like other large-scale replication projects have attempted to do. This means the interpretation of the data is a bit challenging.

Below are suggestions for the authors and readers to consider:

(1) I understand why the authors prefer to mention claims as their primary means of reporting what they found, but it is nested within paper, and that makes it very hard to understand how to interpret these results at times. I also cannot understand at the high-level the relationship between claims and papers. The methods suggest there are 3-4 major claims per paper, but at 400 papers and 1,006 claims, this averages to ~2.5 claims per paper. Can the authors consider describing this relationship better (e.g., distribution of claims and papers) and/or considering presenting the data two ways (primary figures as claims and complimentary supplementary figures with papers as the unit). This will help the reader interpret the data both ways without confusion. I am also curious how the results look when presented both ways (e.g., does shifting to the paper as the unit of analysis shift the figures and interpretation?). This is especially true since the first and last author analysis shows there is varying distribution of papers and claims by authors (and thus the relationship between these is important for the reader).

(2) As mentioned above, I think the biggest weakness is that the authors are taking the literature at face value when assigning if a claim was validated or challenged vs gathering new independent evidence. This means the paper leans more on papers, making it more like a citation analysis vs an independent effort like other large-scale replication projects. I highly recommend the authors state this in their limitations section.

On top of that, I have questions that I could not figure out (though I acknowledge I did not dig super deep into the data to try). The main comment I have is How was verified (and challenged) determined? It seems from the methods it was determined by "Claims were cross-checked with evidence from previous, contemporary and subsequent publications and assigned a verification category". If this is true, and all claims were done this way - are verified claims double counted then? (e.g., an original claim is found by a future claim to be verified - and thus that future claim is also considered to be verified because of the original claim).

Related, did the authors look at the strength of validation or challenged claims? That is, if there is a relationship mapping the authors did for original claims and follow-up claims, I would imagine some claims have deeper (i.e., more) claims that followed up on them vs others. This might be interested to look at as well.

(3) I recommend the authors add sample sizes when not present (e.g., Fig 4C). I also find that the sample sizes are a bit confusing, and I recommend the authors check them and add more explanation when not complete, like they did for Fig 4A. For example, Fig 7B equals to 178 labs (how did more than 156 labs get determined here?), and yet the total number of claims is 996 (opposed to 1,006). Another example, is why does Fig 8B not have all 156 labs accounted for? (related to Fig 8B, I caution on reporting a p value and drawing strong conclusions from this very small sample size - 22 authors). As a last example, Fig 8C has al 156 labs and 1,006 claims - is that expected? I guess it means authors who published before 1995 (as shown in Figure 8A continued to publish after 1995?) in that case, it's all authors? But the text says when they 'set up their lab' after 1995, but how can that be?

(4) Finally, I think it would help if the authors expanded on the limitations generally and potential alternative explanations and/or driving factors. For example, the line "though likely underestimated' is indicated in the discussion about the low rate of challenged claims, it might be useful to call out how publication bias is likely the driver here and thus it needs to be carefully considered in the interpretation of this. Related, I caution the authors on overinterpreting their suggestive evidence. The abstract for example, states claims of what was found in their analysis, when these are suggestive at best, which the authors acknowledge in the paper. But since most people start with the abstract, I worry this is indicating stronger evidence than what the authors actually have.

The authors should be applauded for the monumental effort they put into this project, which does a wonderful job of having experts within a subfield engage their community to understand the connectiveness of the literature and attempt to understand how reliable specific results are and what factors might contribute to them. This project provides a nice blueprint for others to build from as well as leverage the data generated from this subfield, and thus should have an impact in the broader discussion on reproducibility and reliability of research evidence.

Author response:

The following is the authors’ response to the original reviews.

eLife Assessment:

This important study presents an impressive large-scale effort to assess the reproducibility of published findings in the field of Drosophila immunity. The authors analyse 400 papers published between 1959 and 2011, and assess how many of the claims in these papers have been tested in subsequent publications. In a companion article they report the results of experiments to test a subset of the claims that, according to the literature, have not been tested. The present article also explores if various factors related to authors, institutions and journals influence reproducibility in this field. The evidence supporting the claims is solid, but there is considerable scope for strengthening and extending the analysis. The limitations inherent to evaluating reproducibility based on the published literature should also be acknowledged.

We have included in the abstract the limitation of evaluating replicability based on the published literature

Public Reviews:

Reviewer #1 (Public review):

Summary:

The authors set out on the ambitious task of establishing the reproducibility of claims from the Drosophila immunity literature. Starting out from a corpus of 400 articles from 1959 and 2011, the authors sought to determine whether their claims were confirmed or contradicted by previous or subsequent publications. Additionally, they actively sought to replicate a subset of the claims for which no previous replications were available (although this set was not representative of the whole sample, as the authors focused on suspicious and/or easily testable claims). The focus of the article is on inferential reproducibility; thus, methods don't necessarily map exactly to the original ones.

The authors present a large-scale analysis of the individual replication findings, which are presented in a companion article (Westlake et al., 2025. DOI 10.1101/2025.07.07.663442). In their retrospective analysis of reproducibility, the authors find that 61% of the original claims were verified by the literature, 7.5% were partialy verified, and only 6.8% were challenged, with 23.8% having no replication available. This is in stark contrast with the result of their prospective replications, in which only 16% of claims were successfully reproduced. The authors proceed to investigate correlates of replicability, with the most consistent finding being that findings stemming from higher-ranked universities (and possibly from very high impact journals) were more likely to be challenged.

Strengths:

(1) The work presents a large-scale, in-depth analysis of a particular field of science that includes authors with deep domain expertise of the field. This is a rare endeavour to establish the reproducibility of a particular subfield of science, and I'd argue that we need many more of these in different areas.

(2) The project was built on a collaborative basis (https://ReproSci.epfl.ch/), using an online database (https://ReproSci.epfl.ch/), which was used to organize the annotations and comments of the community about the claims. The website remains online and can be a valuable resource to the Drosophila immunity community.

(3) Data and code are shared in the authors' GitHub repository, with a Jupyter notebook available to reproduce the results.

We thank the reviewer for their positive comments.

Main concerns:

(1) Although the authors claim that "Drosophila immunity claims are mostly replicable", this conclusion is strictly based on the retrospective analysis - in which around 84% of the claims for which a published verification attempt was found. This is in very stark contrast with the findings that the authors replicate prospectively, of which only 16% are verified. Although this large discrepancy may be explained by the fact that the authors focused on unchallenged and suspicious claims (which seems to be their preferred explanation), an alternative hypothesis is that there is a large amount of confirmation bias in the Drosophila immunity literature, either because attempts to replicate previous findings tend to reach similar results due to researcher bias, or because results that validate previous findings are more likely to be published.

Both explanations are plausible (and, not being an expert in the field, I'd have a hard time estimating their relative probability), and in the absence of prospective replication of a systematic sample of claims - which could determine whether the replication rate for a random sample of claims is as high as that observed in the literature -, both should be considered in the manuscript.

I still believe that most of ‘verified’ claims are indeed solid, consistent with the feeling of increased knowledge in the Drosophila immunity. Of note, when I started in the field in the early 90, nearly all the articles in Drosophila developmental biology, the dominant field at that time, were solid. I agree that this may represent a past period. Still, perceived irreproducibility is high because affecting claims published in high-impact articles. To address the reviewer’s comments, we have included the possibility of confirmation bias in the discussion of the article.

(2) The fact that the analysis of factors correlating with reproducibility includes both prospective and retrospective replications also leads to the possibility of confusion bias in this analysis. If most of the challenged claims come from the authors' prospective replications, while most of the verified ones come from those that were replicated by the literature, it becomes unclear whether the identified factors are correlated with actual reproducibility of the claims or with the likelihood that a given claim will be tested by other authors and that this replication will be published.

We agree that the ReproSci is a human biased endeavor and may be subjected to bias, notably the one mentioned. Some verified claims might be considered as challenged and some challenged claims as verified. But this is likely not affecting most of our results. Our papers have been posted on a public archive and that the community could scrutinize our data (most authors have checked their claims!). Based on community feedback, we have had to change the status of three claims only: one verified claim became challenged, one challenged claim became unchallenged and one unchallenged became verified. We therefore did not expect major changes in our conclusions. We have provided an estimation of irreproducibility ranging from 10-20% and we believe this is a rather safe interval. This is still significant. The fact that the ReproSci project is a human endeavor, and therefore subject to inherent limitations, is already emphasized in the last section of the discussion.

(3) The methods are very brief for a project of this size, and many of the aspects in determining whether claims were conceptually replicated and how replications were set up are missing.

Some of these - such as the PubMed search string for the publications and a better description of the annotation process - are described in the companion article, but this could be more explicitly stated. Others, however, remain obscure. Statements such as "Claims were cross-checked with evidence from previous, contemporary and subsequent publications and assigned a verification category" summarize a very complex process for which more detail should be given - in particular because what constitutes inferential reproducibility is not a self-evident concept. And although I appreciate that what constitutes a replication is ultimately a case-by-case decision, a general description of the guidelines used by the authors to determine this should be provided. As these processes were done by one author and reviewed by another, it would also be useful to know the agreement rates between them to have a general sense of how reproducible the annotation process might be.

The same gap in methods descriptions holds for the prospective replications. How were labs selected, how were experimental protocols developed, and how was the validity of the experiments as a conceptual replication assessed? I understand that providing the methods for each individual replication is beyond the scope of the article, but a general description of how they were developed would be important.

We have extended the first section of the Result to mention that an extended Material and Methods can be found in the complementary article Westlake et al., eLife 2026 and on the ReproSci website indicated the resources described in the companion articles. Disagreement between the two authors were limited and mostly dealt with textual interpretation (the definition of the major claims, which can affect if they are verified or not). As mentioned above, our articles are now public since more 6 months and we got little request to change the status of a claim (only 3 over more than 1000). Of note many claims from some ancient articles are part of the common knowledge, are verified routinely by many labs in the field, and therefore were easy to categorize.

We also added a sentence to also underline that the scientific value of an article does not fully correlate with its status. A number of problematic articles filled with exaggeration or confused claims were still categorized as verified. Thus, we screen for replicability but they are many other publication issues we could not capture in our study. This point is now mentioned in the discussion.

(4) As far as I could tell, the large-scale analysis of the replication results was not preregistered, and many decisions seem somewhat ad hoc. In particular, the categorization of journals (e.g. low impact, high impact, "trophy") and universities (e.g. top 50, 51-100, 101+) relies on arbitrary thresholds, and it is unclear how much the results are dependent on these decisions, as no sensitivity analyses are provided.

Particularly, for analyses that correlate reproducibility with continuous variable (such as year of publication, impact factor or university ranking, I'd strongly favor using these variables as continuous variables in the analysis (e.g. using logistic regression) rather than performing pairwise comparisons between categories determined by arbitrary cutoffs. This would not only reduce the impact of arbitrary thresholds in the analysis, but would also increase statistical power in the univariate analyses (as the whole sample can be used in at once) and reduce the number of parameters in the multivariate model (as they will be included as a single variable rather than multiple dummy variables when there are more than two categories).

We did not pre-registered our study, which is quite different from other reproducibility projects. We did not know at the start how the project would unfold. Of note, I believe that ‘cleaning’ an extensive field by experts was by itself an achievement and represent another way to assess replicability, with its own limitation.

We agree that the cut-off can be seen as arbitrary. However, publication year has been modeled as a continuous variable through a spline (to allow for non-linearity) in the multivariate analysis. For journal and institutional ranking, we would like to defend and keep the categorical specifications in the main text for two reasons:

- Institution ranking: For institutional ranking, the largest institutional category is “Not Ranked” which we cannot assign a rank and ranks institutions individually only through the top 100. Below that it publishes bands (101-150, 151-200, 201-300, 301400, 401-500) of variable width, which are not straightforward to use continuously without making strong assumptions (e.g midpoint) unsupported by the data.

- Prestige: As mentioned by another reviewer there are career consequences attached to Nature, Science and Cell that are different from other journal, and the effect of moving from a journal from an IF 3 to 8 is different than from 30 to 35. We provide a sensitivity analysis in the supplementary information where we refit the multivariate model with log-transformed impact factor continuous predictor. We obtain very similar results than these in the main text, with the exception that the impact factor of the journal becomes significant. We describe and discuss this results in the supplementary information (section S5), while justifying why we keep the categorical model in the main text.

We are grateful to the reviewer for prompting these additions, which we agree strengthen the paper.

The text added in the SI states:

“As a sensitivity analysis, we refitted the multivariable hierarchical logistic model by replacing journal-impact-factor categories with log₂-transformed continuous impact factor, while retaining all other covariates and random intercepts from the primary analysis. University ranking was retained as a categorical variable because the 2010 Shanghai Academic Ranking reports individual ranks only for institutions ranked 1– 100 and otherwise reports rank bands (101–150, 151–200, 201–300, 301–400, and 401–500); moreover, 441 of 1,006 claims (43.8%) were from unranked institutions, which constitute a distinct group.

Each doubling of journal impact factor was associated with 1.51-fold higher odds that a claim was challenged (OR 1.51, 94% HDI 1.13–2.08). At the mean impact factors of the low-impact and trophy-journal tiers (4.4 and 61, respectively; approximately 3.8 doublings), the continuous model implies a trophy-versus-low-impact odds ratio of approximately 4.8. By contrast, the primary model directly estimated the trophy-versus-low-impact odds ratio as 1.75 (94% HDI 0.59–4.91). Although this wide interval includes 4.8, the point estimates differ between the two models. We interpret the larger implied trophy-journal effect in the continuous model cautiously because it assumes that the log-linear slope estimated predominantly from the many journals with lower impact factors (around 2–12) applies unchanged to the sparsely represented trophy journals (62 claims across Science, Nature, and Cell). We therefore retained the categorical journal analysis as the primary analysis because it directly estimates the trophy-journal comparison using claims published in those journals. Estimates for all other covariates were broadly similar to those in the primary model. In particular, the association with Top 50 university affiliation remained evident (OR 4.04, 94% HDI 1.57–10.08, compared with OR 3.68, 94% HDI 1.50–9.28 in the primary model)”.

[This response is referred later as “Response to categorical variable encoding”]

(5) The multivariate model used to investigate predictors of replicability includes unchallenged claims along with verified ones in the outcome, which seems like an odd decision. If the intention is to analyze which factors are correlated with reproducibility, it would make more sense to remove the unchallenged findings, as these are likely uninformative in this sense. In fact, based on the authors' own replications of unchallenged findings, they may be more likely to belong the "challenged" category than to the "unchallenged" one if they were to be verified.

We thank the reviewer for raising this comment, we agree that the classification of the unchallenged claims as not challenged is difficult given their nature. We refitted the multivariable analysis excluding unchallenged claims, with a classification of Challenged and Unchallenged (Verified, Partially Verified, Mixed). The results are shown in the supplementary section S4 and are extremely similar to the main model.

The text added in the SI states:

“We fit the multivariate model from the main text while excluding unchallenged claims from the analysis. The methods are similar and the convergence and diagnosis checks are good and not reported here for brevity. The model simultaneously adjusts for author demographics, laboratory attributes and journal or institutional prestige and predict a binary outcome for each claim: challenged or not challenged (including Verified, Partially Verified, Mixed). The re-analysis included 659 claims, of which 60 (9.1%) were challenged. The non-challenged reference group comprised of Verified (n=525), Partially Verified (n=64), Mixed (n=10). The Adjusted odds ratios and 94% highest-density intervals are reported in Figure S4. The results are very similar to the analysis with the unchallenged claims in the main text, with only university ranking (Top 50) being a significant predictor (OR 3.34 (1.28–9.01) vs the main text OR 3.68 (94% HDI 1.50– 9.28). The other variable with the biggest changes shows only modest differences:

- High-impact journal OR 1.12 (vs 1.31)

- Trophy journal OR 1.47 (vs 1.75)

- Post-doc OR 1.44 (vs 1.16) and their intervals still include 1. Note that the excluding unchallenged decision condition on follow-up, which is not random and might be influenced by our key variable (journal tier, institutional ranking) and bias results for this model.”

However, we did not replace the main model of the article by this particular model because when you chose only claim that are either verified or challenged (so you exclude unchallenged), the condition that made these claims included in this analysis could be influenced by a cofactor, so our odds ratios might be wrong.

[This response is referred later as “Response to unchallenged exclusion”]

Reviewer #2 (Public review):

Summary:

Lemaitre et al. conducted an analysis of 400 publications in the Drosophila immunity field (1959-2011), performing both univariable and multivariable analyses to identify factors that correlate with or influence the irreproducibility of scientific claims. Some of the findings are unexpected, for instance, neither the career stage of the PI nor that of the first author appears to matter that much, while others, such as the influence of institutional prestige or publication in "trophy journals," are more predictable. The results provide valuable insight into patterns of irreproducibility in academia and may help inform policies to improve research reproducibility in the field.

Strengths:

This study is based on a large, manually curated dataset, complemented by a companion paper (Westlake et al., 2025. DOI 10.1101/2025.07.07.663442) that provides additional details on experimentally documented cases. The statistical methods are appropriate, and the findings are both important and informative. The results are clearly presented and supported by accessible documentation through the ReproSci project.

We thank the reviewer for their assessment.

Weaknesses:

The analysis is limited to a specific field (immunity) and model system (Drosophila). Since biological context may influence reproducibility -- for example, depending on whether mechanisms are more hardwired or variable -- and the model system itself may contribute to these effects (as the authors note), it remains unclear to what extent these findings generalize to other fields or organisms. The authors could expand the discussion to address the potential scope and limitations of the study's generalizability.

We have added a limitation section at the end of the manuscript to discuss these points. We wrote:

“Finally, the findings we report may be specific to a particular field, model organism, and time period, and may be strongly influenced by factors idiosyncratic to their development. Because this analysis is limited to a single field (immunity) and model system (Drosophila), it remains unclear to what extent these findings generalize to other fields, organisms, or more recent studies.”

However, we believe that our study adds on the discussion on reproducibility by providing another approach. Our study also aligns with the feeling that knowledge of Drosophila innate immunity has increased so much in the last decades, and that science is still quite robust in many fields.

Reviewer #3 (Public review):

Summary:

The authors of this paper were trying to identify how reproducible, or not, their subfield (Drosophila immunity) was since its inception over 50 years ago. This required identifying not only the papers, but the specific claims made in the paper, assessing if these claims were followed up in the literature, and if so whether the subsequent papers supported or refuted the original claim. In addition to this large manually curated effort, the authors further investigated some claims that were left unchallenged in the literature by conducting replications themselves. This provided a rich corpus of the subfield that could be investigated into what characteristics influence reproducibility.

Strengths:

A major strength of this study is the focus on a subfield, the detailing of identifying the main, major, and minor claims - which is a very challenging manual task - and then cataloging not only their assessment of if these claims were followed up in the literature, but also what characteristics might be contributing to reproducibility, which also included more manual effort to supplement the data that they were able to extract from the published papers. While this provides a rich dataset for analysis, there is a major weakness with this approach, which is not unique to this study.

Weaknesses:

The main weakness is relying heavily on the published literature as the source for if a claim was determined to be verified or not. There are many documented issues with this stemming from every field of research - such as publication bias, selective reporting, all the way to fraud. It's understandable why the authors took this approach - it is the only way to get at a breadth of the literature - however the flaw with this approach is it takes the literature as a solid ground truth, which it is not. At the same time, it is not reasonable to expect the authors to have conducted independent replications for all of the 400 papers they identified.

However, there is a big difference trying to assess the reproducibility of the literature by using the literature as the 'ground truth' vs doing this independently like other large-scale replication projects have attempted to do. This means the interpretation of the data is a bit challenging.

We acknowledge that our analysis necessarily relies on the published literature; however, two important points should be considered. First, we assessed replicability more than a decade after the original studies, allowing us to evaluate whether their claims have withstood subsequent scrutiny. Second, the evaluation draws on the authors’ strong expertise in the field. Accordingly, the validity of published claims was not accepted uncritically but subjected to informed expert assessment. This is exemplified by the case of the proposed microbicidal role of Duox, which we classified as challenged despite the original article accumulating more than 600 citations.

Overall, our replicability analysis, like similar efforts, relies not only on literature but also on expert judgment. Of course it has inherent limitations, which are now more clearly articulated in the discussion.

Below are suggestions for the authors and readers to consider:

(1) I understand why the authors prefer to mention claims as their primary means of reporting what they found, but it is nested within paper, and that makes it very hard to understand how to interpret these results at times. I also cannot understand at the high-level the relationship between claims and papers. The methods suggest there are 3-4 major claims per paper, but at 400 papers and 1,006 claims, this averages to ~2.5 claims per paper. Can the authors consider describing this relationship better (e.g., distribution of claims and papers) and/or considering presenting the data two ways (primary figures as claims and complimentary supplementary figures with papers as the unit). This will help the reader interpret the data both ways without confusion. I am also curious how the results look when presented both ways (e.g., does shifting to the paper as the unit of analysis shift the figures and interpretation?). This is especially true since the first and last author analysis shows there is varying distribution of papers and claims by authors (and thus the relationship between these is important for the reader).

Articles can contain claims that we found replicable and other that we did not. Thus, we believe than an evaluation by claim is more precise. The extraction of the different claims relies on the reading of the article, but is a human enterprise. There are clearly some unavoidable biases at this step. Of note, the articles analyzed in this study are very different one from another (size, novelty, amount of data…).

However, we agree that the analysis at the article level is interesting, and we followed the reviewer suggestion and provided in Supplementary Section S6 a sensitivity analysis of using article-level effects for the multivariate analysis. We note that because only 43 articles (12.5%) contained at least one challenged claim, this analysis is less robust than the claim analysis. The text added in the SI states:

“Two claims in the same paper might not be independent, so we might want to incorporate an article random effect to the primary Bayesian formulation. However, due to the low number of claim per article (mostly one to four) and only 43 articles (12.5%) contained at least one challenged claim (and most articles contains only one to four claims), we could not manage to fit this model properly (we observe MCMC divergences). As a work-around sensitivity analysis, we repeated the analysis with the article as the unit. Each of 345 articles with complete covariate data (representing 869 major claims, 60 of the 69 challenged claims in the full dataset; the remaining 9 fell in articles excluded for missing covariates) contributed as outcome the proportion of its major claims classified as challenged. The model was a (frequentist) binomial-link fractional-response generalized linear model with HC3 standard errors (the intervals reported here are 94% confidence intervals rather than posterior credible intervals).

Each article contributed equally regardless of its number of claims. The Top-50 institutional effect remained significant under both analysis (OR 5.15, 94% CI 2.18– 12.15; versus a claim-level OR 3.68, 94% HDI 1.50–9.28). In this new model Trophy journals also showed had higher odds of challenged claims than low-impact journals (OR 2.90, 94% CI 1.24–6.78 vs a claim-level OR 1.75, 94% HDI 0.59–4.91). This impact was not significant in the primary model but became significant at using article-level analysis. Other estimates are consistent in direction and magnitude with the primary claim-level model. Model diagnostics were satisfactory.”

Moreover, we added Figure S6 which display visually the relationship between the claims and the article.

[This response is referred later as “Response to claim vs article level analysis”]

(2) As mentioned above, I think the biggest weakness is that the authors are taking the literature at face value when assigning if a claim was validated or challenged vs gathering new independent evidence. This means the paper leans more on papers, making it more like a citation analysis vs an independent effort like other large-scale replication projects. I highly recommend the authors state this in their limitations section.

We have now explicitly acknowledged this limitation in a dedicated section at the end of the revised manuscript. However, as noted above, we do not take the literature at face value. Notably, some of the claims we identify as challenged originate from highly cited articles, which could lead a naïve reader to assume their validity.

Ultimately, this replicabilty project remains a human, expert-driven effort aimed at determining which claims withstand scrutiny within a field. The broader community was invited to engage with and respond to these assessments and we did get some feedbacks. Importantly, claims were not evaluated in isolation but in the context of the entire body of knowledge, and all evaluations were made transparent and explicitly justified in the ReproSci database.

On top of that, I have questions that I could not figure out (though I acknowledge I did not dig super deep into the data to try). The main comment I have is How was verified (and challenged) determined? It seems from the methods it was determined by "Claims were cross-checked with evidence from previous, contemporary and subsequent publications and assigned a verification category". If this is true, and all claims were done this way - are verified claims double counted then? (e.g., an original claim is found by a future claim to be verified - and thus that future claim is also considered to be verified because of the original claim).

When two articles reported the same main claim and this claim was judged as “verified,” each instance was counted in the analysis. However, when validation came from studies published after 2011, the claim was counted only once. Thus, our study specifically evaluates the replicability of claims originating from a defined period (pre-2011).

We note that, in principle, repeated publication of the same finding across many articles could artificially inflate estimates of replicability. However, such cases appeared to be rare in practice, and studies reporting similar claims typically provided complementary evidence—for example, elucidating gene function through either genetic or biochemical approaches.

Related, did the authors look at the strength of validation or challenged claims? That is, if there is a relationship mapping the authors did for original claims and follow-up claims, I would imagine some claims have deeper (i.e., more) claims that followed up on them vs others. This might be interested to look at as well.

We did not perform a systematic mapping between original claims and subsequent follow-up studies. While this is an important and interesting aspect, it was considered beyond the scope of the present work. The current study already represents a substantial investment of time and effort, and the resulting database provides a valuable resource for future analyses. In particular, it could support a dedicated bibliometric study at a later stage.

(3) I recommend the authors add sample sizes when not present (e.g., Fig 4C).

The cited panel appears to correspond to Figure 5C in the revised numbering, as the current Figure 4 contains only panels A and B. We systematically checked the denominators of all figures and revised the captions to report the numbers of claims, authors or laboratories represented, together with the reasons for exclusions. We also added the numbers of authors and claims directly within the author-level scatter plots, with the observational unit explicitly identified.

I also find that the sample sizes are a bit confusing, and I recommend the authors check them and add more explanation when not complete, like they did for Fig 4A. For example, Fig 7B equals to 178 labs (how did more than 156 labs get determined here?), and yet the total number of claims is 996 (opposed to 1,006).

There are 156 labs total, 23 labs contributed publications during both junior and senior stages, and 11 claims lacked a junior/senior classification. This has been explicited in the caption.

Another example, is why does Fig 8B not have all 156 labs accounted for?

The panel does not include all 156 laboratories because we could obtain information about first author training for only 146 PI and laboratories while prior first-author status was unavailable or indeterminate for ten PI. This has been added to the caption.

(Related to Fig 8B, I caution on reporting a p value and drawing strong conclusions from this very small sample size - 22 authors).

The 22 authors are not the total sample size; they constitute the subgroup of PIs who had previously published a first-author paper on Drosophila immunity in another laboratory. In the submitted panel, they were compared with 117 PIs without this training history, giving 139 classified laboratories and 774 claims.

As a last example, Fig 8C has al 156 labs and 1,006 claims - is that expected? I guess it means authors who published before 1995 (as shown in Figure 8A continued to publish after 1995?) in that case, it's all authors? But the text says when they 'set up their lab' after 1995, but how can that be?

Thank you for identifying this inconsistency. Figure 8C intentionally includes the full cohort of 156 PIs and all 1,006 claims. Panel C compares research styles and does not impose a restriction based on year of entry into the field or year of claim publication. Thus, PIs who entered the field before 1995 are included together with all claims attributed to them; the panel is not restricted to claims published after 1995.

The statement that only PIs who established their laboratories after 1995 were included was an erroneous carryover from an earlier version of the analysis. We removed this restriction from the Results and revised the caption to state that exploratory PIs (n = 91) contributed 366 claims, whereas continuity PIs (n = 65) contributed 640 claims.

We also replaced references to the date when PIs “established their laboratory” with “year of entry into the field,” defined as the year of their first Drosophila-immunity publication as either first or last author. During the same check, we revised Panel B to include all 146 PIs eligible for the pre-PI training comparison, contributing 793 claims.

(4) Finally, I think it would help if the authors expanded on the limitations generally and potential alternative explanations and/or driving factors. For example, the line "though likely underestimated' is indicated in the discussion about the low rate of challenged claims, it might be useful to call out how publication bias is likely the driver here and thus it needs to be carefully considered in the interpretation of this. Related, I caution the authors on overinterpreting their suggestive evidence. The abstract for example, states claims of what was found in their analysis, when these are suggestive at best, which the authors acknowledge in the paper. But since most people start with the abstract, I worry this is indicating stronger evidence than what the authors actually have.

We have added a dedicated section on limitations at the end of the Discussion to explicitly acknowledge key constraints, and we also address these points throughout the Discussion where relevant. In addition, these limitations are now in the abstract.

The authors should be applauded for the monumental effort they put into this project, which does a wonderful job of having experts within a subfield engage their community to understand the connectiveness of the literature and attempt to understand how reliable specific results are and what factors might contribute to them. This project provides a nice blueprint for others to build from as well as leverage the data generated from this subfield, and thus should have an impact in the broader discussion on reproducibility and reliability of research evidence.

Thank you for this positive assessment!

Recommendations for the authors:

Reviewer #1 (Recommendations for the authors):

Comments and recommendations are provided in the order that they appear in the manuscript.

Introduction:

- In the second paragraph of the introduction, I don't think all the references refer to "molecular life sciences" (e.g. one of them is from ecology and evolution, for example). I'd also argue that Macleod et al. 2014 refers to risk of bias rather than to reproducibility per se.

We have corrected this in the revised version.

- The author use both "reproducibility" and "replicability", apparently as synonyms, but as definitions for these words vary across sources (e.g. https://arxiv.org/abs/1802.03311, https://www.nationalacademies.org/our-work/reproducibility-and-replicability-in-science, https://osf.io/br9sp/) it may be worth stating their definitions upfront.

Thank you for the references. We agree that the literature contains multiple, sometimes conflicting definitions of these terms, which can lead to confusion. In response to the reviewer’s comment, we have clarified our terminology. Specifically, we now use the term “conceptual replicability” in the abstract and consistently refer to “replicability” and “irreplicability” throughout the manuscript. We hope this revised terminology improves clarity.

- Why are analysis described as "exploratory" and "multivariate"? If it has not been preregistered, shouldn't the multivariate analysis be considered exploratory as well?

Thanks, we agree. The word “exploratory” is confusing as we use it for its structural meaning and not its epistemological meaning. We have replaced it throughout by “descriptive”, as to not imply that the multivariate analysis is preregistered by contrast.

Results:

Methodology and claim assessment

- The methods for the selection, annotation and classification of studies are described very cursorily here, and the Methods section doesn't add that much, but I'll keep these observations to the methods. Still, it's worth mentioning that it's hard to know how reproducible the annotation and classification process is from the information provided.

We have provided more information on the method section on the way to annotate and class claims. Of note, there is more information on the methods on the associated publication Westlake et al., 2025 and on the ReproSci website.

- "Then, experimental work was performed in several laboratories...". Once again, a much better description of how replications were set up and performed is needed.

We have provided information on the number of laboratories (nine) involved in the reproduction of 45 major claims. The laboratories involved in replication are listed in the author list. All the files associated with experimental replication can be found in the supplementary materials of Westlake et al 2025 or the ReproSci website.

- "These claims subjected to experimental validation were selected according to criteria such as appearing suspicious due to the absence of direct follow up or being straightforward to test experimentally". This is rather vague, but if no explicit criteria were set upfront this may be inevitable. Still, it would be useful to know who made this decision.

A first list of claims to be replicated was made by Hannah Westlake and Bruno Lemaitre. The claims that were selected to be replicated were tested by the labs and experimental workers that had the expertise required to test them. The choice of the claims to be tested was not random. Claims that were suspicious are claims that were already tested in the host laboratory without success, claims whose validity are debated at meeting, claims from highranked journal with no follow-up. This is now indicated in the method of the revised version.

Drosophila immunity claims are mostly reproducible

- The percentages given in the initial description of the results pool the results of the retrospective analysis of the literature and the prospective replications performed. As mentioned in the Public Review, I think these are rather distinct ways to assess reproducibility, and would recommend that these two analyses are described separately.

We agree with the reviewer but we believe that we have already well separated the results before and after experimental work in the revised version. This is clearly shown in Table 1. Our text says ‘Importantly, 6.8% of claims (69 out of 1006) were challenged. Among these, 44.9% (31 out of 69) were contested by published articles including 7.2% by the same authors, and 55.1% (38 out of 69) were challenged experimentally as part of the ReproSci project.’

- "These results may reflect the robust scientific standards and methodological rigor of Drosophila research...". Once again, as mentioned in the Public Review, I think publication bias is a plausible explanation, so this should be mentioned as an alternative.

We have mentioned publication bias in the revised version in the section on limitation. But an important point to underline is that this analysis did not take publication results as face fact. This is a critical analysis by experts in the field. As already mentioned, some claims like ‘Duox produces microbicidal ROS” ‘NO is a signaling molecule…’that have been mentioned in multiple articles were considered as challenged.

A second point is that it is likely that publications in some fields are more replicable than in other. Actually, we did not discuss ‘reproducibility issues’ when I started in the field 30 year ago because this was not perceived as an issue. It is normal that reproducibility rate varies according to fields. Nevertheless, we do find important claims with high visibility that are not replicable in our sample. Thus, our findings nuance the ‘reproducibility crisis narrative’ but we still believe that there is concern with reproducibility in life science. Last point, the criterium we have used in our study did not map all the problematic papers, notably some claims might be fundamentally verified but exaggerated in the original paper.

- It would be useful to break down the "partially verified" category somewhere. Illustrative examples of what is meant by "insightful data were accompanied by incomplete interpretations" or "incomplete data were paired with insightful interpretations" would be useful as well.

All the claims are listed in the public database with an explanation as to why they got affected to a category. The categories mixed and partially verified refer to claims that are more complex to categorize.

Here are are two examples from the ReproSci website:

Claim: Relish Rel mutants are very susceptible to bacterial and fungal infection. Partially verified

Assessment: Relish mutants are primarily susceptible to Gram-negative bacterial infections, and do not typically have increased susceptibility to Gram-positive bacteria and fungi which are primarily rebuffed by the Toll pathway, although Imd signaling may provide a minor contribution to defense against these. The high susceptibility of Relish mutant to fungi observed in this article is likely due to the presence of the ebony marker (Lemaitre et al., 1996) [here only one part of the statement has been verified]

Claim: PGRP-SC1a is required for Lys-type peptidoglycan recognition and Toll activation. Mixed

Assessment: A possible role for PGRP-SCs in Toll activity has been strongly debated. Challenged by (Bischoff et al., 2006) who found no effect of PGRP-SC RNAi on activation of Toll (Drs expression) in adult flies (although note that their data show a minor reduction of Drs expression in flies in response to E. faecalis). Supported by (Costechareyre et al., 2016) who used single -SC1 and -SC2 mutants to show that Toll activation (Drs expression) was reduced in -SC1 but particularly -SC2 mutants (~50%) in response to E. faecalis (although survival was not affected). The supplementary data of (Paredes et al., 2011) similarly show that Drs expression was reduced (50%) in response to M. luteus septic infection in PGRP-SCdelta flies (PGRP-SC1A/B -SC2 triple mutant), but concluded that this was due to a secondary effect of Imd overactivation. Note that the phenotype found by (Garver et al., 2006) was much stronger (complete ablation of Drs expression), which is not consistent with any subsequent results and argues for a secondary mutation in this line, although this is not consistent with the successful rescue of Drs expression by transgenic replacement of PGRP-SC1a. Downregulation of Toll in these mutants could be explained if the Toll pathway requires amidase activity of PGRP-SC for immunogenicity (e.g. the sugar backbone without stem peptides is a stronger elicitor than whole peptidoglycan), whereas similar cleavage reduces recognition by the Imd pathway (consistent with the demonstrated requirement of stem peptides for full stimulation of the Imd pathway (Chang et al., 2005; Stenbak et al., 2004)). See annotation for (Mellroth et al., 2003). [here the claim is mostly challenged but there are observations that go in the same directions]

As shown by these examples, category assessment was complex process and relied on human expertise. However, there was very little contestation from the community after the submission of our articles and the opening of the ReproSci website.

A significant fraction of unchallenged claims is non-reproducible

- Excluding the 45 unchallenged major claims that were experimentally tested, we categorized the remaining 240 unchallenged claims into three groups". No information is provided on that classification process (either here or in the methods). Who made these assessments (which seem quite subjective and dependent on field expertise), and how do you know if such judgments are reproducible across different evaluators?

This categorization was done by Hannah Westlake and reviewed by Bruno Lemaitre with few disagreements. This is now indicated in the revised version. The category ‘unchallenged logically consistent” and ‘‘unchallenged logically consistent’ refer to claims that although not directly verified are corroborated or in with current literature. See ReproSci for justifications.

Example from ReproSci website:

Claim: The minimal structure needed to activate the Toll pathway is a muropeptide dimer.

Assesment: Unchallenged logically consistent This is consistent with (Park et al., 2007) who show that on a linearized strand of peptidoglycan, the minimal motif is at least 3 dimers. But it is expected that a non-linearized dimer linked by the peptide bridge could serve as the minimal motif by clustering peptidoglycan.

Claim: eater null flies are susceptible to oral infection with Serratia marcescens Unchallenged logically consistent Inconsistent with the observation that Eater mutants successfully phagocytose Serratia marcescens (Bretscher et al., 2015). This may indicate that another factor affected by eater mutation is required for defense against S. marcescens, such as formation of lamellopodia and filopodia or adhesion of hemocytes to the body wall, or that Eater contributes to binding of Gram-negative bacteria but does not trigger phagocytosis in response to them as it does for Gram-positive bacteria.

Higher representation of challenged claims in trophy journals and from top universities

- Both impact factor and the Shanghai university ranking are continuous variables, but the authors opt to use them as categorical variables in the analysis (i.e. "low-impact", "highimpact", "trophy"; "Top 50", "51-100", "101+"). While there may be legitimate reasons to do this if they feel that these categories are a better descriptor of the underlying reality (e.g. perhaps Cell, Science and Nature are indeed in a category by themselves), it is an unusual decision that leads the comparison to ignore the distinctions between journals/universities within a category. Moreover, it also opens up the opportunity to analysis bias as categories can be set up in many different ways. Thus, if the analysis was not preregistered, I'd recommend that analyses using impact factor and ranking as quantitative variables are added as sensitivity analyses, as these seem to me to be the most natural/less ad hoc way to look at the issue.

See “Response to categorical variable encoding” above.

- As mentioned in the Public Review, the analysis here pools the results from the retrospective analysis based on the literature and the prospective one based on the performed replications. Although this may be justified from a sample size perspective (as the number of challenged claims is not that high), it would be useful to perform this analysis separately on the two sets of results as well, as different trends may be noticed. There are many biases that might come into play here (e.g. results from top institutions being more or less likely to have challenged/unchallenged claims published) and looking at the results separately could help in tearing apart these hypotheses.

The ReproSci project analyzes all the papers of a community during a period of time. As such, the total number of claims is limited and we believe that pooling the data was justified. We provide as sensitivity analysis in Supplement S7, a multivariate model with claims in their pre-experimental classification state:

“As a sensitivity analysis, we refitted the multivariable hierarchical logistic model using only the claim classifications available before experimental validation by the ReproSci project, as the choice of which claims to further validate could have been influenced by covariate (impact factor, university status). The 45 claims tested prospectively were therefore restored to their original “Unchallenged” classification, while all covariates and author-level random intercepts were retained from the primary analysis. The complete-case analysis included 869 claims, of which 28 (3.2%) were challenged. No predictor had a 94% highest-density interval excluding 1, including publication in trophy journals (OR 1.11, 94% HDI 0.33–3.77) and affiliation with a Top50 institution (OR 1.44, 94% HDI 0.48–3.94). Thus, the associations observed in the pooled analysis were not evident when using classifications based exclusively on the retrospective literature review, although estimates were imprecise because of the small number of challenged claims”

The irreproducibility rate has increased over time as the field has grown in popularity

- Again, why not use year as a continuous variable rather than using 5-year windows (as this would effectively include more information)? I think the categorization here is less ad hoc than in the journal/university case, but it is still an unusual decision. More important than this, however, is the fact that pairwise comparisons between periods is probably not an appropriate strategy to look for a time trend, as it excludes all information not related to the pair in each analysis. I'd strongly suggest substituting this for a straightforward regression with publication year as a continuous variable.

We thank the reviewer for this important suggestion. We agree that categorizing publication year into five-year periods may obscure temporal trends.

We therefore modelled publication year continuously using a spline in the multivariable analysis, thereby retaining the full temporal information while allowing for non-linearity. A straightforward linear regression would impose a linear trend, which may not adequately capture changes over time. We retained the five-year groupings solely for visualization.

- "The subsequent increase in unchallenged claims may reflect the rapid conceptual expansion of the field, which likely outpaced the growth in the number of researchers". Do we have data on the growth in the number of researchers in the field? If so, it might be interesting to cite this.

We do not have the number of researchers to assess the growth of the field, however we can see in Figure 8A an increase in the number of laboratories (as identified by last author name) after 1995. This together with the increase in the number of articles (Figure 3A) clearly show the expansion on of the field.

- Isn't a "rise in unchallenged claims" over time expected by chance, as older findings will have more time to have been verified by someone else? I understand that the 14-year window probably mitigates this effect, but it still probably exists in some degree and should be mentioned.

The reviewer is correct that the % unchallenged claims should increase over time, but this does not explain the very low number of unchallenged claims between 1992-2001.

First-author patterns of irreproducibility

- With 69 challenged claims and 289 first authors, it would be impossible to achieve "perfect equality" (e.g. a Gini index of 0.88). To verify how far the observed coefficient deviates from chance, it would be useful to obtain (perhaps via simulations) the Gini coefficient expected by chance given the number of challenged claims/authors, and perhaps derive a p value as well (e.g. the proportion of random permutations in which the index exceeds the observed value).

Very good observation, thank you, we provide this analysis for both first and last author, in the main text and in a supplementary table.

Added first author text:

“The distribution of challenged claims among first authors was highly unequal with a Gini coefficient of 0.881 (inequality index ranging from 0 -perfect equality- to 1 extreme inequality-, Figure 4B): the top 10% of first authors accounted for 73.9% of challenged claims, and the top 20% accounted for all challenged claims. This inequality exceeded that expected by chance (reassigning randomly the 69 challenged labels across individual claims, while preserving each author’s number of claims, p = 0.000001). It also remained greater than expected when claims from the same paper were kept together (p = 0.015; Supplementary Table Sx). The concentration of challenged claims among first authors cannot be explained solely by differences in the number of claims or by clustering within papers.”

Added last author text:

“Challenged claims were also unevenly distributed among leading authors (Gini coefficient = 0.856): the top 10% accounted for 71.0% of challenged claims and the top 20% accounted for 94.2%. The observed inequality exceeded that expected when challenged labels were reassigned across individual claims (p = 0.00081), but not when claims from the same paper were kept together (p = 0.127; Supplementary Table S9). The apparent concentration among leading authors could be explained by multiple challenged claims arising from the same papers.”

Added method text:

“We compared the observed Gini coefficients with two permutation analyses, each based on 1,000,000 permutations. First, we randomly reassigned the challenged labels across individual claims while keeping every claim attached to its original author. This preserved the number of claims contributed by each author. Second, we kept the challenged-claim pattern of each paper together and reassigned these patterns among papers containing the same number of claims. This additionally accounted for the possibility that claims from the same paper were challenged because of a shared problem. For each permutation, we recalculated the Gini coefficient across authors and calculated the p-value as the proportion of simulated coefficients that equalled or exceeded the observed coefficient. Full results are reported in Supplementary.”

- Claims in a single paper may not be fully independent from each other (as both could be irreproducible due to the same error), so it could be worth adding the article as a random variable here.

See answer above: [“Response to claim vs article level analysis”]

- In Fig. 4B, what defines the order of the dots between 0 and 250 (i.e. authors that have no challenged claims)? Is this merely arbitrary? This should be stated more clearly.

Thank you, the order was arbitrary, it has now been fixed by using a secondary sorting (verified claim proportion), and Authors tied on challenged-claim proportion are ordered by verified-claim proportion has been added the legend.

- In Fig. 5C, there are clearly less dots than there are authors. This is likely due to superposition (i.e. there are probably multiple circles with 1 article and 0% verified claims). Nevertheless, it means that the graph conveys an erroneous message. As the x axis is a discrete variable, it may be worth turning it into five categories (e.g. 1 to 5) and adding some jitter to each of them in order to let the reader know how many dots are in each category/%.

Thank you, this was indeed due to superposition. After trying to add a jitter, the graph was still confusing, so we changed the representation to show circle whose size depends on the number of authors. The graph now conveys the size of each block properly.

Lead-author patterns of irreproducibility

- The same comment concerning the Gini index made for the first authors also holds here.

We added a text, see answer above.

- In Fig. 6B, the same comment made for figure 4B also holds.

Thank you, the order was arbitrary, it has now been fixed by using a secondary sorting (verified claim proportion), and Authors tied on challenged-claim proportion are ordered by verified-claim proportion has been added the legend.

- The division between "senior" and "junior" PIs is rather ad hoc here. Once more, why not use "years from first last-author publication" as a continuous variable as a more neutral way to analyze this (as done in Fig. 8)? Also note that it is not clear whether "published a lastauthor article at least five years prior to the considered publication" means any article or one about Drosophila immunity (as in Fig. 8A), and which database was used to examine this (PubMed? Other?).

Senior and junior researchers were classified based on whether they had published a last-author article five years or more using Pubmed (for seniors). We did not specify articles in Drosophila immunity when considering seniority but only having a last author article five year before the publication.

- In Fig. 7C, the same comment made for Fig. 5C also holds, although the solution here is less obvious as there are more possible numbers for "number of articles".

Thank you. As in Fig. 5C, superposition obscured multiple authors occupying the same coordinates. We therefore revised Fig. 7C so that authors with the same number of articles and proportion of challenged claims are represented by a single circle. Circle size and the number shown inside indicate how many authors are represented.

- As far as I could tell, the data on Fig. 8A refers to when authors published their first first/last author papers on Drosophila immunity (which makes it different from Fig. 7, which refers to any article, but I could be mistaken). If this is the case, I'm not sure it's correct to talk about "the period when principal investigators established their laboratories" as mentioned in the text, as (a) they could have published a first author paper in somebody else's lab or (b) they could have started a lab and only later published a paper on Drosophila immunity. "Year of entry in the field" as in the figure legend seems more appropriate.

We have changed for ‘year of entry in the field’ as suggested by the reviewer

- In the same figure, the 1995 cutoff seems completely arbitrary. Why not analyze this as a regression with year of entry as a continuous variable?

The 1995 boundary is historical as it marks the expansion of Drosophila immunity from a marginal subject into a popular field. The expansion of the field at this specific time was probably driven by a combination of factors: major advances in Drosophila genetics and genomic resources and growing knowledge of antimicrobial peptides in the early 1990, a strong interest on innate immunity that drew attention on the power of Drosophila to answer key questions in this new area. These developments attracted new researchers to insect immunity, and the later discoveries of the IMD and Toll pathways in 1995 and 1996 gave the field an additional (and probably even stronger) boost. Our corpus holds 29 articles for 1959– 1994 (0.8 per year) against 371 for 1995–2011 (21.8 per year), and only 13 of the 156 PIs entered before 1995. The comparison is therefore a cohort contrast more than a contrast or search for a cutting point. Moreover, a regression would impose a linear relationship between the different years, which we believe is too constrained on regard of the observed data.

- "We hypothesized that these authors, having gained prior hands-on experience on Drosophila immunity, would be less prone to publishing irreproducible claims.". This is a possibility, but given that most of the sample is retrospective, it's also possible that the field is more prone to publicly challenging findings from newcomers.

Since we do not observe major difference in replicability between senior and junior PI, we still believe that training as first author in a traditional immunity laboratory compared to no training is significant.

Irreproducibility according to research styles

- The objective definition of "continuity" and "exploratory" PIs should be stated here for this to be interpretable (note that this is not clear in the Methods either).

We have better defined how we separate "continuity" and "exploratory" PIs in the methods and in the result section.

Multivariable analysis of predictors of claim irreproducibility

- As mentioned in the Public Review, does it make sense to include "unchallenged" along with "verified" in the outcome? If the objective is to reduce the analysis to a "verified/not verified" claim, wouldn't it make more sense to remove the unchallenged findings (as these are likely uninformative, and based on the authors' own replications may be more related to the "challenged" category than to the "unchallenged" one?

See above, “Response to unchallenged exclusion”

- As stated previously, why not include journal impact factor, university ranking, year of first paper and year of first Drosophila immunity paper, as continuous variables rather than categorical ones (which would effectively lead to a model with less parameters when there are more than 2 categories). Particularly, there is more information to be gained from adding a single variable than from performing individual comparisons between categories when these have a natural order.

See above, “Response to categorical variable encoding”

Discussion:

- As "conceptual reproducibility" is somewhat of a vague concept, it seems important to discuss (both in the Methods and Discussion) how this was operationalized. Although it's obvious some decisions of what constitutes a direct replication will vary on a case-by-case basis, general guidelines on how these criteria were set are needed.

We have defined in the methods used to categorize article claims as well as the link with the other companion article. In the ReproSci database, all the assessment are justified allowing to see how we classified claims.

- "Contrary to the more dramatic narratives...". Here, it is important to emphasize the key differences between this and the cited references: (a) the fact that the majority of the replicability of the sample was found on the basis of a retrospective sample and (b) the fact that the study is dealing with conceptual rather than direct replications. Also, the possibility of confirmation bias should be mentioned as an alternative hypothesis in the last sentence of this paragraph.

This is a good point and we have highlighted that methodological differences between our study and other studies could explain differences in replicability rates.

- "findings that, despite being published, have never been independently tested". A more accurate description may be "have never been independently confirmed or refuted in the published literature", as many (and perhaps most - see Baker 2016) replication attempts may go unpublished, as the following sentences themselves indicate.

We have changed the text according to the reviewer’s suggestion.

- The description of how findings came to be regarded as "suspicious" here is interesting - and an important part of how the sample was determined. In this sense, I'd consider this as part of the methods. Even though defining what makes something "suspicious" may not be completely systematic, a general description of the method that led the researchers to arrive at this list (e.g. the process that seems to be hinted at in this paragraph) deserves a thorough description.

The description of how findings came to be regarded as "suspicious" is now detailed in the methods.

- "Our results suggest the status of a claim being "unchallenged" is not a reliable proxy for its validity". I agree, but this is statement is in direct contradiction with the authors' decision to include unchallenged claims along with validated ones in the binary outcome of the multivariate model.

See above, “Response to unchallenged exclusion”

- "Trophy journals are more likely to publish articles with challenged claims than high or lowimpact, although the difference was not significant." Again, a single analysis using a continuous measure of impact should provide more statistical evidence than the pairwise comparisons.

See above, “Response to categorical variable encoding”

- "However, this higher rate of follow-up work cannot fully explain by itself the higher proportion of challenged claims in trophy journals." Why not, exactly? I don't remember seeing an objective analysis of it.

If we remove the unchallenged claims, we still observe higher rate of irreplicability in trophy journal: 15.85% in trophy journal versus 9% in high-impact and 8.1% in low-impact. So we believe that the higher rate of irreplicability in trophy journals cannot be explained by a lower level of unchallenged.

- The very large paragraph on pages 21-22 could be split into two, one about journals and the other about universities.

This has been done.

- "Our data align with the broader narrative of increasing rates of non-reproducible science.". Does it? The time trend did not seem very clear to me (and I would argue that it was not analyzed properly).

As indicated in the text, we observed an increase in the rate of irreplicable claims over time; however, this trend did not reach statistical significance. We are confident in the robustness of our analysis and therefore did not pursue this question further. While our findings align with concerns about non-replicable science, they also offer a more nuanced perspective, showing that in some fields, replicability remains significantly high.

- "A similar but more acute pattern was observed during the SARS-CoV-2 pandemic." References should be provided here.

We have added this reference to suggest that articles published on SARS-CoV-2 pandemic are overall less reliable than others: An alarming retraction rate for scientific publications on Coronavirus Disease 2019 (COVID-19) Nicole Shu Ling Yeo-The https://doi.org/10.1080/08989621.2020.1782203

- Some of the trends discussed here (e.g. time, continuity vs. exploratory style are not supported (even as a trend) by the multivariate model, and this caveat should be mentioned. An odds ratio of 0.89 with a very wide confidence interval may be too weak in terms of evidence strength to merit a whole paragraph in the introduction discussing it.

We did not mention the question of time horizon or the distinction between continuity and exploratory research styles in the Introduction; however, we believe these issues merit consideration in the Discussion. In particular, the contrast between continuity-driven and exploratory approaches is noteworthy. Although we found no significant difference in the rate of challenged claims between these approaches, there was a significant difference in the rate of unchallenged claims, which are subsequently more likely to be challenged. This finding is important in the current funding landscape, where many agencies prioritize short-term, exploratory projects. Such incentives may inadvertently contribute to issues of irreproducibility.

- There are many other limitations beyond those mentioned, in particular the fact that much of the analysis is retrospective and based on a potentially biased literature. The fact that the analysis does not seem to be preregistered and is dependent on a lot of ad hoc decisions also merits discussion. This should definitely be explored in more detail in the Limitations section.

These limitations are now discussed in the limitation section at the end of the discussion. We agree on the fact that our analysis was not preregistered and that some of our conclusions are raised after analyzing the data. This article is part of a more global project including the databases with all the information. However, full objectivity when analyzing literature is not unachievable, and these critics are inherent to all reproducibility projects.

Methods:

Selected articles, annotation and experimental validation:

- "In brief, a list of 400 publications published before 2011 was generated using a curated search string on the publicly available PubMed database.". Please state that the search string is available in the companion article (e.g. https://doi.org/10.1101/2025.07.07.663442) or include it here.

This is now stated.

- "Selected primary articles were annotated by a single researcher".

By "annotation", do the authors mean the extraction of claims? This is not self-evident. Also, what does the "review" process entails? Do the authors have any data on agreement?

We do not have data on agreement but overall, there were few discrepancies. We agree that the way to section article in separate claims could affect the conclusion in a number of cases. All the data are public and we did not have any request from the community. The fact that this project depends from arbitrage from the two annotators is mentioned in the limitations.

- "Claims were cross-checked with evidence from previous, contemporary and subsequent publications and assigned a verification category."

How was this process performed? This seems to be quite complex and there's hardly any information about it. And once again, do the authors have any data on how reproducible this process would be when performed by different people?

All the data, notably claim assessment and justifications, are available on the website. Findings cross-checking the claim could be identified by knowledge of the authors, checking articles that quotes the articles. This was an enormous amount of work with human arbitrage. We expected more feedback from the community. The reaction of the community will be described in a companion article later.

- The authors mentioned that annotations and verifications were made available for comments on the community, but how were these comments incorporated if different opinions were voiced? Can the authors provide data on how frequent these comments were, and how often they were incorporated?

All data is available on the website, where members of the community can publicly comment on our assessments. The authors can see these comments on the ReproSci website

Table 1:

- Again, the three "partially verified" categories in Table 1 seem to refer to very different situations and it would be interesting to break these down somewhere.

We agree that categories ‘mixed’ and ‘partially verified’ are complex. They represent situations where a decision was not easy. We prefer not to break down those categories in multiple sub-categories.

First and last author classification:

- What does "status of first author" mean?

Status means their position: technician, PhD student, Post-doc, PI….

- "PIs were manually classified as..." - there seems to be something missing in this sentence (e.g. "as senior or junior on the basis of whether they have...")

We have added PIs: PIs were manually classified as senior or junior PIs on the basis on having published an article in Pubmed as last author more than five year ago.

- The distinction between continuity and exploratory is not explained in objective terms. Is there any objective definition of what it means to "continue to work in the field". Publishing a paper in the last X years? And what counts as a "transient" incursion?

We have better explained in the methods this distinction.

Statistical analysis:

- Can the authors define exactly what they mean by "weakly informative priors"?

Weakly informative prior are Bayesian prior distribution that are meant to be vague as to let the data dominate the results (vs the prior belief of the scientist). They can be seen as very close to flat priors, which are not used here because they cause numerical convergence issues.

- The authors mention a lot of variables included in the multivariable analysis, but little information is provided on the categorization. What are "low, high and trophy journals"? What university rankings are used?

We used Shanghai Ranking’s 2010 Academic Ranking of World Universities as our ranking of university. Our binning of ranking in categories make the analysis less susceptible to changes in ranking system. We selected the Shanghai ranking as it primarily evaluates research output and awards (where Times also evaluate teaching, and QS also employer reputation), which we believe represent better the variable than may affect irreplaceability. For the categorization, please see answer above: “Response to categorical variable encoding”

- "Exact formulas (...) are available in the public repository". What repository do authors mean? The project website? The GitHub repository? Please specify and provide a direct link if possible.

We added a link to the main text (it was only in the supplementary document). The Supplementary document contains information on how to reproduce the results.

Reviewer #2 (Recommendations for the authors):

Specific comments:

Some methodological details related to the main conclusions of the paper are missing.

We have extended the methodological section and also better link this article to the companion article.

- Impact factor: It is unclear which year(s) of journal impact factors were used. Trophy journals are defined as those with an impact factor >50 (Science, Nature, and Cell), but according to Figure 2B, there are only three such journals. Are "trophy journals" limited to these three, or are others included that meet the >50 impact factor threshold?

These are the 2022 impact factor, we made that clear in the main text. Moreover, only 3 journals in our datasets had an impact factor >50: Nature, Cell, and Science (so the Trophy Journal category is limited to these 3, but not by definition). This is explicit in Table S1: List of journals with impact factor and claim assessment.

- University ranking: Similarly, please clarify which year(s) of university ranking data were used. Since rankings vary depending on the system (e.g., QS, Times, Shanghai), were the conclusions consistent across multiple ranking sources?

We used Shanghai Ranking’s 2010 Academic Ranking of World Universities as our ranking of university. Our binning of ranking in categories make the analysis less susceptible to changes in ranking system. We selected the Shanghai ranking as it primarily evaluate research output and awards (where Times also evaluate teaching, and QS also employer reputation), which we believe represent better the variable than may affect irreplaceability.

Reviewer #3 (Recommendations for the authors):

Table S4 - for leading author there is a variable of 'historical lab after 1998 continuity' - should that be 1995?

Correct, we changed this label to Trained in Historical laboratory (Comparison restricted to claims published after 1995) to make it more explicit.

Version 2

Abstract

The ReproSci project retrospectively analyzed the conceptual replicability of 1006 claims from 400 papers published between 1959 and 2011 in the field of Drosophila immunity. This project attempts to provide a comprehensive assessment, 14 years later, of the replicability of nearly all publications across an entire scientific community in experimental life sciences. We found that 61% of claims were verified, while only 7% were directly challenged (not replicable), a replicability rate higher than previous assessments. Notably, 24% of claims had never been independently tested and remain unchallenged. We performed experimental validations of a selection of 45 unchallenged claim, that revealed that a significant fraction (38/45) of them is in fact non-replicable. We also found that high-impact journals and top-ranked institutions are more likely to publish challenged claims. In line with the reproducibility crisis narrative, the rates of both challenged and unchallenged claims increased over time, especially as the field gained popularity. We characterized the uneven distribution of irreplicability among first and last authors. Surprisingly, irreplicability rates were similar between PhD students and postdocs, and did not decrease with experience or publication count. However, group leaders, who had prior experience as first authors in another Drosophila immunity team, had lower irreplicability rates, underscoring the importance of early-career training. Finally, authors with a more exploratory, short-term engagement with the field exhibited slightly higher rates of challenged claims and a markedly higher proportion of unchallenged ones. While limited by its reliance on published literature, this systematic, field-wide retrospective study offers meaningful perspective into the ongoing discussion on reproducibility in experimental life sciences.

Introduction

Science predominantly thrives on the continuous aggregation of knowledge, with researchers progressively building upon one another’s discoveries. As results are replicated (when independent studies reach the same conclusions as the original authors of a claim), confidence in the evidence supporting the claim grows. Confidence in publishing results is critical to sustain trust within a scientific community and to allow new research to build on previous findings (Fidler and Wilcox, 2018). Over the past decades, this confidence has been on a decline.

Several recent reports point out that replicability is lower than desirable in molecular life sciences (The Brazilian Reproducibility Initiative et al., 2025; Baker, 2016; DasGupta et al., 2007; Errington et al., 2021b; Prinz et al., 2011). In 2016, an extensive survey indicated that more than 70% of researchers have tried and failed to reproduce another scientist’s experiments, and more than half have failed to reproduce their own experiments (Baker, 2016). Another study suggested that about 85% of biomedical research is ‘wasted’, meaning that inefficiencies and methodological flaws throughout the research process undermine the value of scientific findings (Begley and Ioannidis, 2015; Chalmers and Glasziou, 2009). In addition, 52% of scientists that were surveyed by Nature agree that we are currently facing a significant ‘crisis’ of replicability. Ongoing concerns about replicability in academic research carry important consequences for the private sector. Two pharmaceutical companies, Bayer and Amgen, reported replicability rates of 11% and 25% in two independent efforts to reproduce findings from dozens of groundbreaking basic studies in cancer and related areas (Prinz et al., 2011). Moreover, an extensive project in cancer biology reveals that a very significant fraction of findings reported in high profile journals cannot be replicated or have smaller effect size that initially reported (Errington et al., 2021, 2014). Lack of replicability can have multiple causes, ranging from the effect of cryptic factors not accounted for by authors, misuse of technical tools, spoiled reagents, bias in data reporting, lack of statistical power or even outright fraud (Devezer and Buzbas, 2023; Eisner, 2018; Lesperance and Broderick, 2021; Peng, 2015).

Misleading or erroneous findings can disrupt scientific progress, wasting valuable time and resources as researchers attempt to replicate or use the results of a published article (Stern et al., 2014). In some cases, entire research initiatives may be built upon findings that are later discovered to be inaccurate or flawed (Piller, 2022). This might also result in unfair career outcomes if sensational but irreproducible results are rewarded over rigorous studies, especially in high-impact publications (Lemaitre, 2016; Steneck, 2011). However, full replicability is neither realistic nor necessarily desirable. A certain level of irreplicability may reflect scientific boldness and risk-taking. Cutting-edge explorations of the unknown are expected to involve occasional backtracking as the bigger picture gradually comes into focus. Thus, striving for perfect replicability could paradoxically stifle innovation and promote overly conservative science. Determining the ideal balance is currently an open question (Shiffrin et al., 2018).

Considering the prominent role played by science in contemporary society, there is a growing need to understand how academic incentives influence the reliability of published results. Indeed, the field of metascience - the scientific investigation of science itself - is gaining popularity (Fortunato et al., 2018). Analyzing how replicability affects the scientific enterprise is required to better characterize how science progresses despite mistakes. It can also help to influence policy to improve science replicability, or at least to limit its most detrimental impacts.

Research focusing on the immune system of Drosophila flourished in the early 1990s and is still an active field of investigation. It boasts two thousand articles published since its beginning, dealing with various facets of the immune response of this insect, such as humoral immunity, cellular immunity, response to pathogens, immuno-metabolism, and evolution of the immune system (Westlake et al., 2024). While the field remains modest in size, it gained visibility in the mid 1990s when researchers realized that some mechanisms regulating the Drosophila immune defense are conserved in mammals, including humans. The discovery of Toll-like receptors, stemming from the identification of Toll’s role in the Drosophila immune response (Lemaitre, 2004), established the value and utility of the field to immunological research, a finding that was acknowledged by a Nobel prize awarded to Jules Hoffmann in 2011.

The medium size of this research field and the lack of direct biomedical influence are ideal for a metascience analysis. The Drosophila immunity field is substantially smaller than other fields of research such as HIV, microbiota or cancer, where the rapid accumulation of articles forbids any comprehensive evaluation of scientific claims by an individual or a laboratory. With fewer direct translational or biomedical influences, we expect scientists working on Drosophila immunity to be less affected by external biases, and to prioritize recognition by peers. Moreover, Drosophila research is generally perceived as solid, because it relies predominantly on its powerful genetics and well-controlled experimental conditions, so the field is particularly interesting for this analysis.

In the framework of the ReproSci project, we retrospectively analyzed the replication of 400 articles published between 1959 and 2011 in the field of Drosophila immunity research-that is, nearly all the major articles published by a community of scientists in experimental life science over 50 years (Westlake et al., 2026). For each article, we identified the main, major, and minor scientific conclusions, hereafter referred to as “claims”. The 14-year interval between the time the latest article was published and our study, during which scientific knowledge has advanced and other laboratories have had the opportunity to attempt replication, provides the necessary perspective to efficiently assess the validity of the claims. In a first step, we created a publicly accessible website, displaying all the information on the articles, the claims, and their validity (https://ReproSci.epfl.ch/). Second, we contacted the Drosophila scientific community to provide feedback via this website on the replicability of the claims. In a third step, several teams attempted to replicate 45 major and 11 minor claims that remained unchallenged in the literature at the time. The present article focuses on the meta-science findings, while the scientific consequences of the ReproSci project in the Drosophila immunity field are presented in a companion article (Westlake et al., 2026).

This article analyzes the pattern of replicability with two objectives: first, to comprehend the level of replicability within the experimental life science community, and second, to examine whether instances of non-replicability tend to aggregate among journals, first authors, and lead researchers. Such clustering may identify individual characteristics and environmental factors that influence science replicability. We open the Results section with a detailed descriptive analysis of the effect of individual categorical variables on reproducibility outcomes. Then, to assess the joint influence of all covariates and account for the clustering of multiple claims by the same author, we fit a multivariate model to account for confounding effects across our covariates. Because challenged claims are rare and the overall sample is modest, statistical power remains limited. However, taken together, i) the descriptive and ii) multivariate analyses form a single narrative, the former describing the observed patterns exhaustively, and the latter testing the robustness of the associations.

Results

Methodology and claim assessment

A list of 400 publications published before 2011 was generated using a curated search string on the publicly available PubMed database (Westlake et al., 2026). This limitation was imposed to allow a time frame in which follow-up studies may be published that may verify, expand on or contradict the content of the publications selected for the study. Scientific claims were divided into three categories. The main (one per article), major (3-4) and minor (2-4) claims were manually extracted from 400 articles in the field of Drosophila immunity published before 2011. We then assessed the claims of each article to determine if they were supported or refuted by subsequent publications. The claims were manually categorized by H.W. and B.L. into four broad groups: i) verified by subsequent studies (when later studies confirmed the claim), ii) challenged (when later studies contradicted or disproved the claim), iii) unchallenged (when no later study addressed or attempted to replicate the claim), and iv) partially verified or mixed (when later studies did not lead to a definitive conclusion (see Table 1 in Methods for classification and see (Westlake et al., 2026) and ReproSci website (https://reprosci.epfl.ch/) for extended description of the project). Then, experimental work was performed in several laboratories to test the validity of 45 major and 11 minor unchallenged claims using alternate methods. These claims subjected to experimental validation were selected according to criteria such as appearing suspicious due to the absence of direct follow up or being straightforward to test experimentally. Our data and findings were put on a community-accessible website that opened to the public in early July 2023. The present article focusses on the pattern of irreplicability and statistical analyses were performed on the 1006 major claims only (referred to as claims below) of the database as updated on December 20th 2024. In this article, challenged claims will interchangeably be called irreplicable. Our project focuses on assessing the strength of the claims themselves (i.e. “indirect reproducibility” or “conceptual replicability”) rather than testing whether the original methods produce repeatable results (“direct reproducibility”). Thus, our conclusions do not directly challenge the initial results leading to a claim, but rather the general applicability (i.e. generalizability) of the claim itself (Goodman et al., 2016; Meng, 2020).

Drosophila immunity claims are mostly replicable

After an extensive literature review to assess the validity of claims published from 1959 to 2011, combined with experimental testing of 45 major claims, we found that approximately 61% of major claims (610 out of 1006 across the 400 selected articles) were fully verified (Table 2). Among these 610 verified claims, 91.6% were corroborated by independent laboratories, 7.2% were confirmed by the same laboratory, and 1.1% were verified through experimental work conducted as part of the ReproSci project. This verification rate is higher than that observed in other fields where reproducibility has been systematically assessed (The Brazilian Reproducibility Initiative et al., 2025; Amaral et al., 2019; Da Costa et al., 2022; Errington et al., 2021; Fidler et al., 2017; Youyou et al., 2023). These results may reflect the robust scientific standards and methodological rigor of Drosophila research, which benefits from the availability of diverse and powerful genetics approaches and from a collaborative culture of rapid reagent sharing, that facilitates replication across laboratories.

We also observed that 23.8% of major claims had no published follow-up (Table 2, Figure 1); these unchallenged claims are particularly interesting and will be discussed in detail in the next section. Additionally, 7.5% of the claims were categorized as partially verified. This category encompasses cases where insightful data were accompanied by incomplete interpretations, or where incomplete data were paired with insightful interpretations. Importantly, 6.8% of claims (69 out of 1006) were challenged. Among these, 44.9% (31 out of 69) were contested by published articles including 7.2% by the same authors, and 55.1% (38 out of 69) were challenged experimentally as part of the ReproSci project. Thus, the experimental validation performed during this project doubled the number of reported challenged claims. Finally, 1.2% of major claims were classified as “mixed” (Table 2). This category includes claims that presented particularly challenging issues (e.g., conflicting data), making it impossible to definitively categorize them as either verified or challenged.

Flow of the study.

Claims were first classified in five categories based on literature. 45 unchallenged claims were selected to be experimentally tested and classified either as challenged or as verified. We further classified some unchallenged as consistent or inconsistent depending on their consistency with current knowledge.

In conclusion, contrary to the prevailing “reproducibility crisis” narrative, a significant proportion of the findings published in the field of Drosophila immunity before 2011 has been subsequently verified by independent studies.

[utbl1]

A significant fraction of unchallenged claims is non-reproducible

Our analysis highlights a significant proportion of “unchallenged” claims that were not followed by subsequent studies and whose validity was unknown prior to our experimental evaluation. Most reproducibility studies have focused on verified or challenged claims so little is known about the vast number of statements generated in a scientific community without subsequent direct validation.

Despite the lack of validation, these “unchallenged claims” are often assumed to be valid and contribute to the broader construction of scientific knowledge. An important challenge is to estimate the percentage of unchallenged major claims that are actually non-replicable.

Prior to our experimental validation, we identified a total of 285 unchallenged claims, representing 28.3% of all major claims published up to 2011. We conducted experimental evaluations of 45 unchallenged major claims with the assistance of multiple laboratories (Westlake et al., 2026). The selection of unchallenged claims for testing was intentionally biased toward two categories: (i) those that appeared suspicious in light of the existing literature (31 claims), and (ii) those that were relatively straightforward to test experimentally (14 claims). Among the 45 unchallenged major claims tested, 84% (38 out of 45) could not be experimentally replicated and were then reclassified as “challenged by the ReproSci project.” Conversely, only 16% (7 out of 45) were successfully replicated and reclassified as “verified by the ReproSci project” (Figure 1). Notably, all 31 suspicious unchallenged claims were challenged, while 7 out of the 14 unchallenged “easy to test” claims were also challenged. Excluding the 45 unchallenged major claims that were experimentally tested, we categorized the remaining 240 unchallenged claims into three groups: 1) unchallenged logically consistent – claims not directly tested but consistent with current knowledge of the field (111 claims); 2) unchallenged logically inconsistent – claims that appear inconsistent with current knowledge (22 claims); 3) Unclassified – claims that could not be definitively categorized (107 claims) (Figure 1). In the remainder of the text, we analyze the claims following ReproSci experimental validation, labelling the 45 selected claims as either challenged or verified.

This analysis reveals that a substantial number of claims remain unchallenged even a decade after publication. Moreover, our findings suggest that a considerable proportion of these claims may not be replicable.

Higher representation of challenged claims in trophy journals

Articles published in high-impact journals receive high visibility, a key factor in career progression and funding security (van Wesel, 2016). Additionally, high-impact publications are increasingly used as metrics to evaluate university rankings. Consequently, scientists face significant internal and external pressures to publish in such prestigious journals (Udesky, 2025). On the other hand, high-impact journals aim to attract the most surprising findings, which are inherently more likely to be challenged than more conventional studies. Based on this, we hypothesized that articles published in high-impact journals are more likely to include non-replicable results. To test this hypothesis, we estimated the number of challenged claims published in journals categorized as having low or medium-impact factors (<10), high-impact factors (>10 and <50), and “trophy journals” of the highest reputation (>50, which corresponds to Science, Nature, and Cell only) using the 2022 impact factors. The results are summarized in Figure 2, with a more detailed breakdown by journal provided in Table S1. Compared with low-impact journals, where 5.8 % of major claims were challenged, the proportion rose to 7.8 % in high-impact journals (odds ratio [OR] = 1.36, 95 % confidence interval [CI] 0.80–2.32) and more than doubled to 12.9 % in trophy journals (OR = 2.39, 95 % CI 1.06–5.40). We recall that the OR expresses how much more (OR > 1) or less (OR < 1) likely a claim is to be challenged relative to articles published in low-impact journals, while the 95 % CI indicates the range of values compatible with random sampling error; OR intervals that exclude 1 denote a statistically significant effect. Thus, our analysis shows that major claims published in low-impact journals are significantly more likely to be replicable than major claims published in trophy journals. One explanation for this trend is that claims published in high-impact journals are more frequently subjected to replication attempts or used as a foundation for further research, thereby increasing the likelihood of identifying irreplicable results. Indeed, low-impact journals include a higher share of claims remaining unchallenged (29.3 %, 95 % CI 25.9–32.9 %) compared to high-impact journals (13.9 %, 95 % CI 10.5–18.2 %) and to trophy journals (17.7 %, 95 % CI 10.2–29.0 %). However, variation in the number of unchallenged claims does not explain the higher proportion of challenged claims in trophy journals compared to high-impact journals. If we take into consideration the percentage of verified, unchallenged and challenged claims, Figure 2B shows that high-impact but not trophy or low-impact journals contain the highest proportion of verified claims.

Percentage of claims by category according to journal-impact and institution ranking.

(A, B) The distribution of claims across the five assessed categories for articles published in low-impact, high-impact, and trophy journals. The number of claims is indicated at the top. Panel A presents raw counts, while Panel B shows normalized data. (C) The distribution of claims across the five assessed categories according to the ranking of the last author institution in Shanghai Ranking’s 2010 Academic Ranking of World Universities. The number of claims is indicated at the top, while the number of institutions in each category is indicated at the bottom. The Not Ranked categories encompass research institutes (ex. CNRS), hospitals and other institution that are not included in the Shanghai ranking. In this figure, and in the rest of the paper, the number nc indicates the number of claims in each category, while the number at the bottom highlight the count in each category (e.g there are 11 high-impact journals, and 73 Top-50 institutions).

Thus, we conclude that while trophy journals may indeed attract some of the most impactful science, they also appear more likely to include irreplicable claims.

Higher representation of challenged claims from top universities

We then investigated whether the percentage of irreplicability varies among laboratories according to the prestige of their affiliate universities. To test this hypothesis, we classified all the major claims published according to the Shanghai ranking of the last author university. Figure 2C shows that claims from laboratories at top-50 universities were challenged far more often 15.8 % (95 % CI 11.4–21.6 %) than other institutions, up to a four-fold increase in the odds of being questioned (OR = 4.69, 95 % CI 2.53–8.70) versus 3.9 % for unranked institutions, which notably include hospitals and research institutes such as the French CNRS. Even after adjusting for journal tier, see multivariate model below, university prestige remains the dominant signal: claims from top-50 institutions are 3.6 times more likely to be challenged than those from unranked universities (OR 3.68, 94 % HDI 1.50–9.28). By comparison, publishing in a “trophy” journal adds only a notable and uncertain risk (OR 1.75 0.59–4.91), while other journal or ranking bands show no clear effect. Put simply, elite affiliation more than journal prestige best predicts irreplicability in our sample.

The irreplicability rate has increased over time as the field has grown in popularity

The reproducibility crisis narrative suggests that irreplicability has increased over recent decades, posing a significant challenge to the life sciences (Baker, 2016; Hunter, 2017). Our dataset spans 50 years, from the onset of Drosophila immunity research as a relatively marginal field to its raise to prominence, with a notable increase in publications since the mid-1990s driven by growing interest in innate immunity. We analyzed fluctuations in irreplicability over time by grouping major claim assessments into time windows: pre-1991, followed by 5-year intervals until 2011. The percentage of irreplicability was calculated for each time window (Figure 3). Across the five windows examined, the proportion of challenged claims climbed steadily from 0% in the small corpus prior to 1991 (0/29 claims), to 4.6% during 1992–1996, and then further increased to 7.7% between 1997 and 2001. This upward trend peaked at 8.3% during the 2002–2006 period, before slightly declining to 6.1% between 2007 and 2011 (Figure 3). Despite these apparent temporal shifts, none of the pairwise comparisons between periods reached statistical significance, with p-values consistently ≥ 0.15.

Evolution of claim replicability over time.

(A, B) The distribution of claims by categories of assessment over time is shown in A, and the proportion normalized over the total number of claims per time period is shown in B.

Thus, while the trends highlighted in our study are consistent with a raise of irreplicability, we cannot exclude that the observed upward trajectory might reflect sampling variation rather than a definitive sustained temporal trend. Further studies focusing on articles published after 2011 could clarify whether the mild decrease we observed marks the beginning of a downward trend or reflects fluctuations at a consistently high level.

Interestingly, we observed a peak in verified claims between 1992 and 2001, accompanied by a lower percentage of unchallenged claims (Figure 3). This period coincided with the early expansion of the field when many researchers worked on similar research questions with a relatively narrow conceptual framework of innate immunity. The subsequent increase in unchallenged claims may reflect the rapid conceptual expansion of the field, which likely outpaced the growth in the number of researchers, reducing opportunities for overlapping or validating research efforts. An alternative explanation for this trend is that in the early stages of the field, researchers uncovered the largest, easily detectable effects. As the field matured, focus shifted to subtler phenotypes and complex regulatory mechanisms, making replication more challenging.

Overall, our time-course analysis suggests that irreplicability has increased over the decades, aligning with the reproducibility crisis narrative. However, an alternative hypothesis is that irreplicability varied alongside the field’s popularity, peaking once when the field became popular and had attracted newcomers who may not have been fully acquainted with its established standards or who opportunistically joined the field. Furthermore, the rise in the percentage of challenged claims was accompanied by an increase in unchallenged claims, indicating a reduced degree of validation within the field.

First-author patterns of irreplicability

Having analyzed the pattern of irreplicability across a large number of articles, representing contributions from multiple laboratories over five decades, we next investigated the distribution of irreplicability among authors and laboratories. A central question is whether irreplicability is evenly distributed across the scientific community, or if hubs — specific authors or laboratories —are responsible for producing more invalid claims. First, we explored whether replicability is associated with the first author, who typically conducts the work, or with the last author, who manages the laboratory and may foster a research environment that produces irreplicable results.

To explore the relationship between irreplicability and first authorship, we computed the percentage of “challenged” articles attributed to each first author. Additional information on first author was obtained through interviews with principal investigators (PIs) or via public LinkedIn profiles, including variables such as gender and position in the laboratory. Among the 400 articles analyzed by the ReproSci project, additional information for a first author was identified for 394 articles, corresponding to 289 unique first authors and 991 major claims. These authors contributed an average of 3.43 claims each. Notably, 46 authors (15.9%, 95% CI: 12.2–20.6%) published at least one non-replicable claim. Figure 4A displays the distribution of challenged claims among first authors. The distribution of challenged claims among first authors was highly unequal with a Gini coefficient of 0.881 (inequality index ranging from 0-perfect equality- to 1-extreme inequality-, Figure 4B): the top 10% of first authors accounted for 73.9% of challenged claims, and the top 20% accounted for all challenged claims. This inequality exceeded that expected by chance (reassigning randomly the 69 challenged labels across individual claims, while preserving each author’s number of claims, p = 0.000001). It also remained greater than expected when claims from the same paper were kept together (p = 0.015; Supplementary S9). The concentration of challenged claims among first authors cannot be explained solely by differences in the number of claims or by clustering within papers. We next investigated whether irreplicability varied by the gender of the first author. The proportions of challenged articles were similar for male and female first authors: 7.3% and 6.7%, respectively, and this difference was not statistically significant (OR = 1.09, 95% CI: 0.66–1.78, Fisher’s exact p = 0.802; Figure 5A).

Distribution of irreplicability according to first authors.

A) Distribution of irreplicable claims among first author. We note that all irreplicable claims are produced by less than 20% of the first authors. Both panel includes 289 first authors contributing 991 claims from 394 articles. Six articles containing 15 claims were excluded because the first author was not identified. B) Distribution of irreplicability for all first authors. Each dot represents an author, ranked according to the proportion of challenged claims. The size of the circle indicates the number of claims and the color the percentage of verified claims. Authors tied on challenged claim proportion are ordered by verified claim proportion.

Influence of parameters associated with first author on replicability.

A) Repartition of claims in each assessment category according to the gender of the first authors. The gender of four authors, contributing 10 claims, could not be inferred based on first name ambiguity explaining why the total number of authors is 285. B) Repartition of claims in each assessment category according to status of the first authors. The stages of career of 279 first authors were obtained by interviewing last authors, querying Linkedln public profiles or extrapolated according to the publication list. Senior staff refers to permanent position scientist working in a lab. ND stands for Not Determined. C) Proportion of challenged claims as a function of the number of articles published by each of the 289 identified first authors. Coincident observations are combined into a circle. Circle area and the number shown indicate the number of first authors represented. Color indicates the percentage of claims classified as verified.

We were able to determine the career stage of 279 of the 289 identified first authors at the time of the article was published, most being PhD students (n=168) or postdoctoral researchers (n=87), with smaller numbers identified as technicians (n=5), Masters students (n=5), PIs (n=8), or permanent staff (n=6) (Figure 5B). No major differences in irreplicability were observed between PhD students and post-docs. While we did observe a higher proportion of challenged claims among PIs and Master’s students, and a lower proportion among technicians and permanent staff, these trends are difficult to interpret due to small sample sizes. Finally, we tested the relation between the number of articles published by a first author and the proportion of irreplicable research. No significant trend was observed (linear regression; r2=0.01, p=0.088; Figure 5C), suggesting the level of author productivity is independent of irreplicability. In summary, we observed considerable variability in irreplicability among first authors, suggesting that individual authors may substantially contribute to reproducibility issues.

Lead authors pattern of irreplicability

We then turned to the question of whether irreplicability varies according to laboratory. As it is common practice in biology, we assigned the last author (referred to as PI for principal investigator) to represent the laboratory. The 400 analyzed articles were co-authored by 156 leading authors from various institutions, predominantly from France, Sweden, Japan, South Korea, and the USA. For each last author, we estimated the number of major claims in each of the five categories. The variation in irreplicability across these PIs is shown in Figure 6A. A statistical dispersion analysis in Figure 6B revealed that challenged claims were also unevenly distributed among leading authors (Gini coefficient = 0.856): the top 10% accounted for 71.0% of challenged claims and the top 20% accounted for 94.2%. The observed inequality exceeded that expected when challenged labels were reassigned randomly across individual claims (p = 0.00081), but not when claims from the same paper were kept together (p = 0.127; Supplementary S9). The apparent concentration among leading authors could be explained by multiple challenged claims arising from the same papers. We next explored whether there were differences in irreplicability between male and female last authors. The percentage of challenged claims was slightly higher for male last authors (7.0%, 95% CI: 5.5%– 9.0%) compared to female last authors (5.8%, 95% CI: 3.1%–10.7%), but this difference was not statistically significant (OR = 1.22, 95% CI: 0.59–2.51; Fisher’s exact p = 0.73; Figure 7A).

Distribution of irreplicability according to last authors (PI).

A) Distribution of irreplicable claims among leading authors. Both panels include all 156 last authors, representing 1,006 claims from 400 articles. B) Distribution of irreplicability for all leading authors. Each dot represents a leading author, ranked according to the proportion of challenged claims. The size of the circle indicates the number of claims and the color the percentage of verified claims. Authors tied on challenged claim proportion are ordered by verified claim proportion.

Influence of various parameters associated with the last author on replicability.

A) Distribution of claims by category according to the gender of all 156 last authors. B) Distribution of claims by category according to the seniority of the last authors. Senior PI were defined as having published a last author article at least 5 years before the considered articles. The total of junior and senior authors is higher than the total number of last authors as 23 PIs published both as junior and senior. 11 claims from four articles are excluded because they lacked a junior/senior classification. C) Proportion of challenged claims as a function of article count for all 156 last authors. Authors with identical coordinates are combined into a single circle; circle size and internal labels indicate the number of leading authors represented.

Pressure to publish is typically stronger among junior PIs who often lack tenure and must establish their laboratory. To examine whether this pressure might be associated with irreplicability, we classified PIs as junior or senior based on whether they had published a last-author article at least five years prior to the considered publication. Although Figure 7B indicates a slightly higher proportion of challenged claims among junior PIs (7.2%, 95% CI: 5.1%–10.2%) compared to senior PIs (6.5%, 95% CI: 4.8%–8.8%), this difference was not statistically significant (OR = 1.12, 95% CI: 0.68–1.83, Fisher’s exact p = 0.70). This modest difference was further attenuated when considering mixed and partially verified claims.

We then explored whether the percentage of irreplicability tended to decline according to the number of published articles as last author (Figure 7C). We could hypothesize that PIs that have transiently work on Drosophila immunity would have more irreplicable claims, or on the contrary that the most productive laboratories compromise speed over rigor. However, no significant trend was observed between the percentage of challenged claims and the number of articles published by an author (linear regression; r2=0.001, p=0.763).

We found that the percentage of irreplicability has increased over time as the field became more popular. We therefore explored if this increase in irreplicability affects all PIs or only new PIs that joined the field at a later time. We show in Figure 8A the proportion of challenged claims by last authors according to the publication date of their first article on Drosophila immunity, either as last or as first authors. We observed a significant association between irreplicability and the period when principal investigators (PIs) entered the field. Specifically, PIs, who entered the field on 1995 or later, had a significantly higher percentage of challenged claims (8.1%, 95% CI: 6.4%–10.2%) compared to those who entered the field before 1995 (2.6%, 95% CI: 1.2%–5.6%; OR = 3.24, 95% CI: 1.38–7.59, Fisher’s exact p < 0.003). Among the PIs working in the Drosophila immunity field, we identified 22 who previously published at least one first-author article in this field during their PhD or postdoctoral training in another laboratory. We hypothesized that these authors, having gained prior hands-on experience on Drosophila immunity, would be less prone to publishing irreplicable claims. Consistent with this hypothesis, we observed lower levels of irreplicability among these experienced authors (3.4%, 95% CI: 1.8%–6.4%) compared to authors who had not been trained in laboratories working on Drosophila immunity (8.7%, 95% CI: 6.6%–11.4%; OR = 2.67, 95% CI: 1.28–5.54, Fisher’s exact p = 0.007; Figure 8B).

Patterns of irreplicability by last author according to year of entry in the field and research style.

A. The proportion of challenged claims by last author is shown according to the time at which the author published their first paper as a first or last authors in the field. It includes all 156 PIs and 1,006 claims; 13 PIs entered the field before 1995 and contributed 227 claims, including 6 challenged claims; 143 entered in 1995 or later and contributed 779 claims, including 63 challenged claims. The increase in irreplicability is associated with PIs whose first publication in the Drosophila immunity field, as either a first or last author, occurred in 1995 or later. Circle size indicates the number of published articles by each author. B. Distribution of claims by last author according to whether they have published at least one first author article on Drosophila immunity in another laboratory. Prior-experience status was available for 146 PIs. Ten PIs whose first-author publications occurred after they had already become PIs were excluded from comparison. C. Distribution of claims according to research style for all 156 PIs. Proportion of challenged claims by PIs who have continuously work on the Drosophila immunity (‘continuity PIs’) or made a transient incursion into the field (‘exploratory PIs’).

We conclude that irreplicability varies significantly across laboratories with a limited number of laboratories accounting for most of the irreplicable claims. As for first authors, author gender is not significantly associated with reproducibility. Our studies point to a significant role of experience: PIs that had previously worked in another Drosophila immunity laboratory to the extent that they published a first author article in that laboratory were less prone to publish irreplicable claims.

Irreplicability according to research styles

Among the various career progression strategies, one involves deepening expertise and specialization within the same field, while another entails transitioning to different fields or exploring different career paths. This may reflect the personality of the scientist, either opting for a more detailed, focused approach or seeking continual new challenges. Scientists from the past were often spending their life on a specific topic within a field. In the last decades, changing field has become the norm (Zeng et al., 2019). Accordingly, scientists are not funded by institutions to work long-term on a single topic anymore, but rather funding is granted by agencies for specific short-term projects. Project grants define research with the primary goal to do ‘breakthrough discoveries’, not accumulative research, which is sometimes treated as the ‘pedestrian’ or the ‘stamp collector’ way of doing research. This leads scientists to move quickly into new research fields in search for opportunities of funding and recognition, a nuanced hypothesis supported by the recent description of a “pivot penalty” for scientists that change career tracks (Hill et al., 2025). We can define these two strategies as either “continuity PIs,” which spend their career on one core topic, and “exploratory PIs”, which regularly venture into new topics. To further explore this question, we classified all last authors as “continuity” PIs or “exploratory” PIs. Exploratory PIs are defined as those who made a transient incursion in the Drosophila immunity field before moving on to another field or leaving academia. In contrast continuity PIs worked continuously in the Drosophila immunity field throughout their careers. We observed that continuity PIs had a lower proportion of challenged claims (6.1%, 95% CI: 4.5%–8.2%) compared to exploratory PIs (8.2%, 95% CI: 5.8%–11.5%). However, this difference was not statistically significant (OR = 1.38, 95% CI: 0.84–2.26, Fisher’s exact p = 0.24; Figure 8C). We also noted a markedly higher share of unchallenged claims among exploratory PIs than among continuity PIs, 32.2 % (95 % CI 27.7%–37.2%) versus 19.1 % (95 % CI 16.2%–22.3%), respectively. This difference was statistically significant (OR = 2.02, 95 % CI 1.50– 2.71; Fisher’s exact p < 0.001), reinforcing the idea that researchers who move on to new topics tend to leave a larger fraction of their earlier findings without follow-up or validation by the community. Since unchallenged claims are more likely to be irreplicable, this suggests that irreplicability is higher in the work of PIs with an exploratory style, as opposed to those engaged in a continuity-driven research style.

Multivariable analysis of predictors of claim irreplicability

We have presented above a comprehensive descriptive data analysis, examining one variable at a time. However, drawing robust conclusions necessitates the joint analysis of all covariates, while appropriately accounting for the non-independence of claims made by the same authors. Thus, we fitted a multivariate Bayesian mixed-effects logistic model to uncover the effect of each covariate, while taking into account all other confounding variables at our disposal (Abril-Pla et al., 2023; Capretto et al., 2022)(Figure 9). The model included 869 claims from 345 articles, contributed by 251 first authors and 140 last authors; the remaining 137 claims, including 9 of the 69 challenged claims, were excluded because information was unavailable for at least one variable used in the model.The model simultaneously adjusts for author demographics, laboratory attributes and journal or institutional prestige and predict a binary outcome for each claim: challenged or not challenged (includings Verified, Partially Verified, Mixed and Unchallenged) (See in Supplementary S3 for a similar model to predict unchallenged or not unchallenged claims, and Supplementary S4, S5, S6 and S7 for sensitivity analysis, respectively excluding unchallenged claims, encoding the impact factor continuously, introducing an article-level effect and using claims status before the validation by the ReproSci project). As the claims from the same first or last author are not independent, we included a mixed effect for each first author (251) and each last author (140). These effects capture the influence of a particular author in driving a claim reproducibility. We conducted extensive test on model formulation and convergence (see methods, and Supplementary information for posterior predictive and diagnostics plot). As this analysis is Bayesian, we report the Highest Density Interval (HDI) instead of the confidence interval. We observe that none of the first-author factors (gender: OR= 0.94, 94% HDI: 0.41–2.11; trainee stage: OR = 1.16, 94% HDI: 0.49–2.77) or leading-author factors (gender, seniority, continuity, or having entered the field before 1995) showed a decisive association with irreplicability once every other covariate was taken into account (all HDIs span 1). We still observe a trend toward reduction of challenged claims for lead authors having being trained in Drosophila immunity or having entered the field before 1995. Figure S3 shows that lead author with an exploratory life style tends to have significantly more unchallenged claims (OR: 2.22, 1.18, 4.55). In contrast, university ranking and journal level covariates dominated the model: articles from top-50 universities were almost four times more likely to contain a challenged claim than those from un-ranked institutions (OR 3.68, 1.50–9.28), and papers in trophy journals (Science, Nature, Cell) carried an elevated uncertain risk (OR 1.75, 0.59–4.91) compared to low-impact journals. Intermediate effects for high-impact journals and mid-ranked universities were positive but imprecise. Publication year was modelled as a continuous variable using a three-knot B-spline to allow for non-linear time trends. None of the individual spline coefficients differed credibly from zero, indicating that once journal tier and the other covariates were included the residual calendar-year effect on irreplicability was negligible, indicating that time of publication itself is not a major factor.

Multivariable analysis of predictors of claim irreplicability.

The forest plot reports the odds ratio of each covariate effect on a claim becoming challenged. The multivariate analysis included 869 claims from 345 articles, including 60 challenged claims, contributed by 251 first authors and 140 last authors. Points report adjusted odds ratios and horizontal lines show 94% highest-density intervals.

The hierarchical component comprised 251 random intercepts for first authors and 140 for PIs. The posterior distribution of these intercepts (see Supplementary Information) is broad, confirming that some individuals or laboratories have a systematically higher or lower propensity for challenged claims. Nevertheless, only the three most extreme first-author effects had 95 % highest-density intervals that excluded zero (and all with positive irreplicability) meaning that statistically robust clustering of irreplicability is confined to a handful of contributors; for the vast majority of authors the collected evidence does not support significant differences in replicability from what is expected from a random distribution.

Taken together, the model confirms our descriptive data analysis, namely that it is the prestige of institution and publication in trophy journals rather than the personal characteristics of the scientists that best predicts which claims will later prove to be irreplicable.

Discussion

The ReproSci project presents an in-depth analysis of reproducibility in an experimental science field with a 14-year retrospective period. This initiative focuses specifically on conceptual replicability —the ability to independently reproduce key findings based on the information provided— rather than direct reproducibility using the exact same data and conditions (Plesser, 2018). Replicability is of critical importance to the scientific community who seeks to build upon published claims. One of our project’s key strengths lies in its comprehensive scope: it spans over fifty years of research, from 1959 to 2011, within a specific scientific community, resulting in a robust database that synthesizes the main findings across decades. This time frame encompasses the formative years of the field, characterized by a small number of pioneering laboratories, followed by a period of rapid expansion driven by increased interest, heightened competition, and publication in high-impact journals. This growth phase was marked by a proliferation of knowledge and conceptual innovation, with novel approaches to studying the immune system emerging. With the awarding of the Nobel Prize in 2011 and subsequently the shift toward more translational research, the growth pace of the Drosophila immunity field has since stabilized, and the field remains highly active. Although our study centers on a specific experimental domain at a particular moment in time with a specific geographical development, it reveals broader patterns of irreplicability that are likely generalizable. These findings contribute meaningfully to the ongoing conversation about the reproducibility crisis in the experimental life sciences (Fanelli, 2018; França and Monserrat, 2018; Hunter, 2017; Peng, 2015).

Contrary to the more dramatic narratives surrounding the reproducibility crisis (The Brazilian Reproducibility Initiative et al., 2025; Baker, 2015; Errington et al., 2021), our study shows that a substantial proportion of claims —61%— can be considered as verified. The proportion of claims that have been directly challenged remains relatively low (7%), though likely underestimated. This indicates that research in the Drosophila immunity field is rather solid, possibly due to the power of Drosophila genetics, the collective ethos of sharing within the community, a less competitive environment, and a lack of influence associated with translational research. Since our study is largely based on published literature, we cannot exclude the possibility that the high replicability rate we report is inflated by confirmation bias in the Drosophila immunity field. This bias could arise either because attempts to replicate previous findings tend to converge on similar results due to researcher expectations, or because studies confirming earlier results are more likely to be published (Nissen et al., 2016; Song et al., 2013). However, we consider this explanation unlikely, as the validity of published claims was not accepted uncritically but subjected to informed expert assessment who did not take claims at face value when assessing whether they were validated or challenged. Notably, we find that even highly cited articles—often perceived as particularly robust—contain claims that have later been challenged. However, the categories we are using in this project do not capture all issues related to replicability, but the conceptual validation of the claims. This difference might explain the higher level of replicability we observed. Overall, our findings are consistent with the view that knowledge of Drosophila innate immunity has expanded substantially over recent decades, while also suggesting that scientific findings in this field remain broadly robust.

A novel and important contribution of our work is the identification of a large fraction of unchallenged claims —findings that have never been independently confirmed or refuted in the published literature. The presence of these unchallenged claims could result from the rapid expansion of the field, limiting the possibility of overlapping research that would validate them. However, the existence of unchallenged claims could also be explained if they are not replicable. For instance, the absence of follow-up studies could reflect authors that are active in the field being aware of the fragility of specific unchallenged claims, and so not pursuing them. The existence of unchallenged claims could also stem from the reluctance of other researchers to publish contradictory statement due to the stigma of publishing negative results, the so-called “file drawer” effect (Rosenthal, 1979). Indeed, many scientists report abandoning verification efforts when unable to reproduce a colleague’s findings, fearing personal conflict or professional retaliation (DeCoursey, 2006; Kannan and Gowri, 2014; Nissen et al., 2016). Our experimental validation demonstrates that a significant proportion of unchallenged claims are, in fact, non-replicable. However, we did not randomly select the unchallenged claims that were to be experimentally tested, and we favored suspicious claims and claims that were easy to experimentally test. Thus, the true rate of reproducibility of unchallenged claims should not be directly extrapolated from our experimental validation attempts. However, we found that none of unchallenged, suspicious claims could be replicated. While confirmation bias driven by our initial suspicion cannot be entirely excluded, the absence of follow-up in the published literature for over a decade later supports our interpretation that these findings were not robust. Of note, we considered many of these unchallenged claims were considered suspicious based on informal exchanges within research communities, that reported difficulties in replicating the findings. Most of these non-replicable suspicious claims were published in high-impact or trophy journals (26/31) with great visibility, perhaps contributing to this privately-shared stigma within the community. Taken together, our findings support the notion that “big stories” published without subsequent follow-up may be indicative of irreplicability.

Another takeaway from our experimental validation is that approximately 50% of the unchallenged claims were not previously flagged as problematic failed the experimental validation test regardless. While a claim could remain unchallenged for years because it is valid and no one has anything novel to add to it, our results suggest the status of a claim being “unchallenged”is not a reliable proxy for its validity. If we reclassify the dataset by counting logically consistent unchallenged claims as “verified” and logically inconsistent ones as challenged, and conservatively assume that 50% of the remaining unchallenged claims fall into the latter, we estimate that the true proportion of challenged claims may approach 14%. When we include mixed and partially verified claims, a rough upper-bound estimate suggests that at least 80% of all published claims on Drosophila immunity between 1959-2011 are verified or likely to be true leaving 20% of claims either challenged or difficult to interpret. Importantly, challenged claims can include work that was done well, but where the field has advanced in its understanding and can better interpret the original findings.We believe the percentage of challenged claims likely varies across scientific fields, depending on intrinsic challenges such as the complexity of biological processes, methodological variability, and the field’s attractiveness, which may increase competitive pressure. It is also important to emphasize that our study, given this specific categorization, does not account for articles that—despite containing validated claims—could still be considered problematic due to exaggeration or to the use of fragile methodologies.

In a recent editorial, Fiala and Diamendis suggested that reproducibility efforts should focus on high-impact papers as these articles are most deleterious to the scientific community (Fiala and Diamandis, 2018). In our database, we indeed found that “trophy journals”, considered to contain the most impactful science, are more likely to publish articles with challenged claims than high or low-impact, although the difference was not significant. Three intertwined factors can explain why high-impact factor journals have higher rate of non-reproducibility. First, the uncertainty that comes with exploring the frontiers of knowledge may lead to investigations that are founded on an incomplete understanding of the major factors influencing the results. Second, the unconventional nature of some high-impact discoveries, and the fact that claims reported in these journals generally attract more attention, increases the likelihood that other researchers will attempt to replicate them — thereby raising the number of challenged claims. And third, systemic biases that pressure scientists to publish in these journals for prestige might encourage scientists to take shortcuts in finalizing their studies, increasing the chance of challenged claims. In our study, we observed a decreased proportion of unchallenged claims in articles published in high-impact and trophy journals indicating that claims published in these journals are more replicable. However, this higher rate of follow-up work cannot fully explain by itself the higher proportion of challenged claims in trophy journals. Compared to low rank and trophy journals, high-ranking journals have the best ratio of verified to challenged claims, as low-impact journals have a high percentage of unchallenged claims. Strikingly, we observed higher rate of non-replicable claims for scientists working in universities in the top 50 Shanghai ranking. Previous work has suggested journal rank as a major associate of irreplicability (Brembs et al., 2013), and that high-impact journals favour authors from leading institutions (Kulal et al., 2025). Our study highlights how university ranking is among the most striking correlates of irreplicability. Taken together, the cause and effect relationship at play here may be driven most by factors associated with institutional prestige, such as trust among networks of colleagues that act as editors and reviewers, leading to greater acceptance rates at high-impact journals. Importantly, our results suggest this greater acceptance does not strictly reflect a high quality of the work being published, but rather a willingness to trust irreplicable work more frequently if it comes from a prestigious institute. Our study spans a period during which most research was done in North America, Europe, Japan and South Korea and does not capture the rise of science in China, or rapid growth of scientific output across the global north and south, where scientists in some countries face strong incentive for ‘high impact’ publications (Delgado-López-Cózar et al., 2021; Quan et al., 2017). The fact that the Drosophila immunity field monitored in this study developed initially in Sweden and France, which have unique research community characteristics, might affect some our conclusions or limit them to a specific period. Future studies in other areas of research might help to confirm if working in top ranked universities correlates with higher rate of non-replicable science.

Our data align with the broader narrative of increasing rates of non-replicable science. One possible explanation is a general decline in the standards of rigor in recent years in the life sciences; however, a non-mutually exclusive hypothesis is that the rise in irreplicability may reflect the increasing popularity of certain fields, creating greater variance by attracting more researchers, who themselves may not be fully acquainted with the field’s established norms and methodological rigor, or who move opportunistically from field to field to publish striking papers. When research on the gut microbiota gained momentum in the 1990s, it was rapidly positioned as a promising avenue to explain a wide array of diseases. Substantial funding and media attention enabled the swift development of the field. However, many early studies lacked methodological rigor and could not be reproduced (Wesolowska-Andersen et al., 2014). The lower scientific standards tolerated in this emerging area —compared to more established fields— facilitated the publication of high-profile articles, encouraging a rapid influx of new research groups. In this context, scientific novelty and perceived importance often took precedence over precision and reproducibility. A similar but more acute pattern was observed during the SARS-CoV-2 pandemic (Ioannidis et al., 2022; Yeo-Teh and Tang, 2021). Alternatively, this reproducibility crisis might be more linked to specific features of our time and the competitiveness of science, that may lead some scientist to distorting science to favor their own success (Lemaitre, 2016).

Importantly, our analysis shows that irreplicability is not evenly distributed across all researchers. While irreplicability can stem from the inherent difficulty or novelty of certain research questions, our findings point to the significant influence of both the first author (who performs the work) and the laboratory environment. Surprisingly, we did not observe a major effect of experience level: irreplicability rates were similar between PhD students and postdoctoral researchers and did not markedly decrease with the number of publications. Similarly, we only observe a slightly but not significant higher irreplicability rate among junior faculty compared to senior professors. This suggests that irreplicability is not inherent to early career stage but rather to certain research practices and environments. That said, we did observe a higher irreplicability rate among PIs that entered the field from a non-Drosophila immunity background. This observation supports the value of traditional scientific training pathways, where early-career researchers are mentored in well-established environments before gaining independence.

Another noteworthy finding is our suggested link between irreplicability and “continuous” or “exploratory” research style. We observed that authors with a more exploratory research style have slightly higher proportions of challenged claims and a significantly higher proportion of unchallenged claims, the latter being more likely to be challenged as shown by our experimental replication. Exploratory researchers have worked only transiently in the field, often leaving it after its peak period of popularity. In recent years funding agencies and universities favored this exploratory research style focusing on breakthroughs, notably by funding projects that may be rapidly discontinued. As a counterpoint of this, our study highlights the value and reliability of laboratories that tend to maintain long-term commitment to a specific field.

An important aspiration in the field of metascience study would be the ability to predict claims or articles that are likely irreplicable. Our retrospective analysis identifies several criteria that correlate with irreplicability, although not all of them reach significance. These include i) articles being published by authors working in top-ranked institutions, ii) articles being published in trophy journals, iii) the absence of follow-up studies (i.e. being ‘unchallenged’), and iv) articles being published by authors with an exploratory research style.In contrast, having published as first author in another laboratory within the field, and publishing in high-impact journals but not trophy journals correlate with higher replicability. The absence of follow-up studies regarding claims published in high-impact journals is likely the most useful proxy to question the replicability of these claims.

Limitations

This work has several limitations. First, the definition of the claim, the validity, the determination of the validation was influenced by the annotating authors (HW and BL). While the ReproSci website was opened to the community, many PIs were reluctant or not interested in the project of testing old claims that represent “scientific cold cases”. Second the leading author is stakeholder of the project having published several articles that were included in the analysis, including 44 articles with B.L. as co-author with 4 as first and 21 as last author. Thus, this may not represent the general view of the community. Finally, the findings we report may be specific to a particular field, model organism, and time period, and may be strongly influenced by factors idiosyncratic to their development. Because this analysis is limited to a single field (immunity) and model system (Drosophila), it remains unclear to what extent these findings generalize to other fields, organisms, or more recent studies. However, we believe that despite these limitations the systematic and retrospective analysis of a research field provides insights to better apprehend how science progresses despite errors and the impact of irreplicability.

Methods

Selected articles, annotation, experimental validation of the ReproSci project

We provide here a succinct method of the ReproSci reproducibility project (see the companion article (Westlake et al., 2026) and the ReproSci website for more details). In brief, a list of 400 publications published before 2011 was generated using a curated search string on the publicly available PubMed database. This limitation was imposed to allow a time frame in which follow-up studies may be published that may verify, expand on or contradict the content of the publications selected for the study. Selected primary articles were annotated by a single researcher (HW) and reviewed by a second researcher (BL) to ensure consistency and accuracy across the project. An online database, ReproSci (https://ReproSci.epfl.ch/), was constructed to organize the annotations and provide an interactive and searchable interface. Claims were extracted through careful reading of the selected primary articles and organized according to their importance within the paper: a single Main claim representing the title claim of the paper or overarching finding; 3-4 Major claims emphasized in the abstract or those contributing to the overarching finding; and a variable number of Minor claims representing other findings of note not emphasized in the abstract. Claims were cross-checked with evidence from previous, contemporary and subsequent publications and assigned a verification category (see Table 1). A short statement explaining the evidence and citing sources was written to justify the categorization and provide additional context for the claim; this assessment was linked to each claim and can be viewed on the ReproSci website. The category ‘unchallenged logically consistent” and ‘‘unchallenged logically inconsistent’ refer to claims that although not directly verified are corroborated or not in with current literature (see ReproSci for justification) When two articles reported the same main claim that was judged as “verified,” each instance was counted in the analysis. However, when validation came from studies published after 2011, the claim was counted only once. Thus, our study specifically evaluates the reproducibility of claims originating from a defined period (pre-2011). A subset of unchallenged claims (45 major and 11 minor) was selected for direct experimental reproduction by various authors (see author list) from nine laboratories with relevant expertise. Unchallenged claims that were considered as ‘suspicious’ are claims that were already tested in the host laboratory without success, claims whose validity are debated at meeting, claims from high-ranked journal with no follow-up. This project is concerned with the accuracy of the conclusions themselves rather than the reliability of methods, so we employed conceptual replicability or indirect or inferential reproducibility, where experimental procedures that may be different from those originally employed are used to independently verify the claim. Thus, an article’s scientific value does not necessarily align with whether it has been verified or challenged, and our categorical framework does not capture all issues related to reproducibility. The ReproSci website was made available to the Drosophila scientific community on July 11, 2023, with an email to encourage to comment on the annotations and verifications and contribute evidence.

First authors and last author classification

Information regarding the gender and status of first authors were either obtained by interviewing PIs (for 191 first authors) or found on publicly available websites (LinkedIn and university websites) or were implied based on their first name usage (gender) and their PubMed publication lists (for the other 100 first authors). When some of these information could not be retrieved, these authors were not included in the analysis.

PIs were manually classified as senior or junior whether they have or have not published a last author article in PubMed at least five years before the considered publication. Continuity and exploratory PIs were manually classified accordingto their PubMed publication list. Continuity PIs are PIs that have continued to work in the Drosophila immunity field until now or to the end of their scientific career (deceased, retired, left academia). Exploratory PIs are PIs that have done a transient incursion in the Drosophila immunity field, moving to other topics of research.

Statistical analysis

The analysis was done on the 1006 major claims of the 400 articles. The cleaned database contains one row per major claim with linked article-, author- and laboratory-level covariates, which was subsequently analyzed. Every output, counts, proportions and graphics, exploratory summaries and Bayesian mixed-effects models are defined under a reproducible Python workflow, which enable the full replication of the main results and every figure in the manuscript.

The ninety-five-percent confidence intervals (CI) for single proportions were determined using Wilson’s score method. Whenever a binary predictor (e.g. author gender) was cross-classified with the binary outcome “claim challenged”, we reported the odds ratio with its 95% CI and the two-sided p-value from Fisher’s exact test; The raw contingency tables for every binary predictor appear in the Supplementary document.

We compared the observed Gini coefficients with two permutation analyses, each based on 1,000,000 permutations. First, we randomly reassigned the challenged labels across individual claims while keeping every claim attached to its original author. This preserved the number of claims contributed by each author. Second, we kept the challenged-claim pattern of each paper together and reassigned these patterns among papers containing the same number of claims. This additionally accounted for the possibility that claims from the same paper were challenged because of a shared problem. For each permutation, we recalculated the Gini coefficient across authors and calculated the p-value as the proportion of simulated coefficients that equalled or exceeded the observed coefficient. Full results are reported in Supplementary Document.

Multivariable modelling

Two Bayesian regressions were fitted with weakly-informative priors. Posterior summaries are presented as odds ratios (OR) with 94 % highest-density intervals (HDI).

Hierarchical logistic regression for claim irreplicability

The outcome was whether a major claim was later classified as challenged. Fixed-effect predictors captured article prestige, author characteristics and laboratory context: journal tier (Low, High, Trophy), university-ranking tier (Not ranked, 101 +, 51–100, Top 50), publication year modelled with a three-knot B-spline, first-author gender and training stage (PhD vs post-doc), last-author gender, laboratory seniority (junior vs senior PI), continuity style (continuous vs exploratory) and whether the laboratory had published in the field before 1995. Claims from one author not independent, so random intercepts were included for both first-author key and last-author key.

All models were estimated with the dynamic-HMC NUTS sampler (target _accept = 0.9); four independent chains of 4,000 tuning iterations and 4,000 draws were run for every model. Convergence was confirmed both visually and numerically (all Rhat≤1.01; all effective sample sizes > 400). Supplementary Fig. 1 posterior-predictive checks; Exact formulas, priors, sampling statements and notebooks are available in the public repository (Supplementary Document).